---
title: "Analysis of co-releases of chemicals in endpoint groups"
author: "Paige Varner"
date: "2024-09-13"
output: html_document
---


#load libraries
```{r}
library("dplyr")
library("tidyr")
library("bigrquery")
library("DBI")
library("pheatmap")
library("RColorBrewer")
library("Rmisc")
library("stats")
library("rstatix")
library("ggplot2")
detach(package:Rmisc)
detach(package:plyr)

install.packages("tidyverse")

endpoints = read.csv("~/TSCA/CRA/CRA_Prioritization/Data/TSCA_endpoints.csv")

```


#query data from BQ (dataframe with TRI ID, facility name, chemicals, and total sum of on site releases for the 5 year span)
```{r}
	# set up database connection
	con <- dbConnect(
	  bigrquery::bigquery(),
	  project = "edf-aq-data",
	  dataset = "tri",
)

SQL = "	
	SELECT
	trifd,
	facility_name,
	chemical,
	SUM(on_site_release_total) as sum_onsite_releases_2017_2022
	FROM `edf-aq-data.tri.tri_basic_2017-2022` 
	GROUP BY trifd, facility_name, chemical"

tri = dbGetQuery(con,SQL)

tri = filter(tri, sum_onsite_releases_2017_2022 > 0)
```



#fixing multiple names for same chemical (need to add up releases for chemicals of the same name after fixing the names for individual trifds)
```{r}
#fix chemical names
tri$chemical = gsub("o-Xylene", "Xylene", tri$chemical)
tri$chemical = gsub("p-Xylene", "Xylene", tri$chemical)
tri$chemical = gsub("m-Xylene", "Xylene", tri$chemical)
tri$chemical = gsub("Xylene (Mixed Isomers)", "Xylene", tri$chemical)

tri$chemical = gsub("Hydrogen cyanide", "Cyanide Compounds", tri$chemical)

tri$chemical = gsub("Nickel compounds", "Nickel", tri$chemical)

tri$chemical = gsub("Arsenic compounds", "Arsenic", tri$chemical)

tri$chemical = gsub("Cadmium compounds", "Cadmium", tri$chemical)

tri$chemical = gsub("Chromium compounds", "Chromium", tri$chemical)

tri$chemical = gsub("Lead compounds", "Lead", tri$chemical)

tri$chemical = gsub("Cobalt compounds", "Cobalt", tri$chemical)

tri$chemical = gsub("Nonylphenol Ethoxylates", "Nonylphenol", tri$chemical)

tri$chemical = gsub("Barium compounds", "Barium", tri$chemical)


#add up chemicals with same name for each trifd to get total releases of those chemicals 
tri <- tri %>%
  group_by(trifd, chemical) %>%
  summarise(onsite_releases_2017_2022 = sum(sum_onsite_releases_2017_2022, na.rm = TRUE))

```

#add endpoint designations to tri data, filter based on endpoints
```{r}
tri_endpoints <- tri %>%
  left_join(endpoints, by = c("chemical" = "tri_name")) %>%
  filter(!is.na(analysis_name))

carc <- tri_endpoints %>%
  filter(carc %in% c("Known", "Probable"))

cns <- tri_endpoints %>%
  filter(cns %in% "yes")

cardioresp = tri_endpoints %>%
  filter(cardio == "yes" | resp == "yes")

devrep = tri_endpoints %>%
  filter(dev == "yes" | repro == "yes")

lkt = tri_endpoints %>%
  filter(lkt %in% "yes")
```


#calculate percent co-releases overall for each group: # of coreleases/unique facilities where # of coreleases is total number of facilities - unique facilities
```{r}
#cns = 39%
((length(cns[["trifd"]]) - length(unique(cns[["trifd"]]))) / length(unique(cns[["trifd"]])))*100

#carc = 42%
((length(carc[["trifd"]]) - length(unique(carc[["trifd"]]))) / length(unique(carc[["trifd"]])))*100

#cardioresp = 63%
((length(cardioresp[["trifd"]]) - length(unique(cardioresp[["trifd"]]))) / length(unique(cardioresp[["trifd"]])))*100

#lkt = 43%
((length(lkt[["trifd"]]) - length(unique(lkt[["trifd"]]))) / length(unique(lkt[["trifd"]])))*100

#devrep = 20%
((length(devrep[["trifd"]]) - length(unique(devrep[["trifd"]]))) / length(unique(devrep[["trifd"]])))*100
```



# Function to count unique trifd for a given pair of chemicals
```{r}
count_unique_trifd <- function(data, chem1, chem2) {
  data %>%
    filter(chemical %in% c(chem1, chem2)) %>%
    group_by(trifd) %>%
    filter(n_distinct(chemical) == 2) %>%  # Ensure both chemicals are present
    ungroup() %>%  # Ungroup to avoid errors in summarisation
    summarise(unique_trifd_count = n_distinct(trifd)) %>%  # Count unique trifd
    pull(unique_trifd_count)  # Extract the count as a numeric value
}
```


#co-releases in carc
```{r}
# Get all unique chemicals
chemicals <- unique(carc$chemical)

# Get all possible pairs of chemicals
chemical_pairs <- combn(chemicals, 2, simplify = FALSE)

# Loop through each pair and get the count of unique trifd
results <- lapply(chemical_pairs, function(pair) {
  count <- count_unique_trifd(carc, pair[1], pair[2])
  data.frame(chemical1 = pair[1], chemical2 = pair[2], unique_trifd_count = count)
})

# Combine the results into a single dataframe
carc_co <- do.call(rbind, results)

#make dataframe of number of facilities releasing each chemical
unique_carc <- carc %>%
  group_by(chemical) %>%
  summarise(total_release = n_distinct(trifd)) 

# Join dataframes to get total_release for chemical1 and chemical2
carc_co <- carc_co %>%
  left_join(unique_carc, by = c("chemical1" = "chemical")) %>%
  rename(total_release1 = total_release)

carc_co <- carc_co %>%
  left_join(unique_carc, by = c("chemical2" = "chemical")) %>%
  rename(total_release2 = total_release)

# Calculate percent corelease for heatmaps 
carc_co <- carc_co %>%
  mutate(perc_co = 100 * ((2 * unique_trifd_count) / (total_release1 + total_release2)))

# save data in csv
write.csv(carc_co, file = "~/TSCA/CRA/CRA_Prioritization/Output/carc_corelease.csv")
```


#co-releases in cns
```{r}
# Get all unique chemicals
chemicals <- unique(cns$chemical)

# Get all possible pairs of chemicals
chemical_pairs <- combn(chemicals, 2, simplify = FALSE)

# Loop through each pair and get the count of unique trifd
results <- lapply(chemical_pairs, function(pair) {
  count <- count_unique_trifd(cns, pair[1], pair[2])
  data.frame(chemical1 = pair[1], chemical2 = pair[2], unique_trifd_count = count)
})

# Combine the results into a single dataframe
cns_co <- do.call(rbind, results)

#make dataframe of number of facilities releasing each chemical
unique_cns <- cns %>%
  group_by(chemical) %>%
  summarise(total_release = n_distinct(trifd)) 

# Join dataframes to get total_release for chemical1 and chemical2
cns_co <- cns_co %>%
  left_join(unique_cns, by = c("chemical1" = "chemical")) %>%
  rename(total_release1 = total_release)

cns_co <- cns_co %>%
  left_join(unique_cns, by = c("chemical2" = "chemical")) %>%
  rename(total_release2 = total_release)

# Calculate percent corelease for heatmaps 
cns_co <- cns_co %>%
  mutate(perc_co = 100 * ((2 * unique_trifd_count) / (total_release1 + total_release2)))

# save data in csv
write.csv(cns_co, file = "~/TSCA/CRA/CRA_Prioritization/Output/cns_corelease.csv")

```


#co-releases in cardioresp
```{r}
# Get all unique chemicals
chemicals <- unique(cardioresp$chemical)

# Get all possible pairs of chemicals
chemical_pairs <- combn(chemicals, 2, simplify = FALSE)

# Loop through each pair and get the count of unique trifd
results <- lapply(chemical_pairs, function(pair) {
  count <- count_unique_trifd(cardioresp, pair[1], pair[2])
  data.frame(chemical1 = pair[1], chemical2 = pair[2], unique_trifd_count = count)
})

# Combine the results into a single dataframe
cardioresp_co <- do.call(rbind, results)

#make dataframe of number of facilities releasing each chemical
unique_cardioresp <- cardioresp %>%
  group_by(chemical) %>%
  summarise(total_release = n_distinct(trifd)) 

# Join dataframes to get total_release for chemical1 and chemical2
cardioresp_co <- cardioresp_co %>%
  left_join(unique_cardioresp, by = c("chemical1" = "chemical")) %>%
  rename(total_release1 = total_release)

cardioresp_co <- cardioresp_co %>%
  left_join(unique_cardioresp, by = c("chemical2" = "chemical")) %>%
  rename(total_release2 = total_release)

# Calculate percent corelease for heatmaps 
cardioresp_co <- cardioresp_co %>%
  mutate(perc_co = 100 * ((2 * unique_trifd_count) / (total_release1 + total_release2)))

# save data in csv
write.csv(cardioresp_co, file = "~/TSCA/CRA/CRA_Prioritization/Output/cardioresp_corelease.csv")

```


#co-releases in devrep
```{r}
# Get all unique chemicals
chemicals <- unique(devrep$chemical)

# Get all possible pairs of chemicals
chemical_pairs <- combn(chemicals, 2, simplify = FALSE)

# Loop through each pair and get the count of unique trifd
results <- lapply(chemical_pairs, function(pair) {
  count <- count_unique_trifd(devrep, pair[1], pair[2])
  data.frame(chemical1 = pair[1], chemical2 = pair[2], unique_trifd_count = count)
})

# Combine the results into a single dataframe
devrep_co <- do.call(rbind, results)

#make dataframe of number of facilities releasing each chemical
unique_devrep <- devrep %>%
  group_by(chemical) %>%
  summarise(total_release = n_distinct(trifd)) 

# Join dataframes to get total_release for chemical1 and chemical2
devrep_co <- devrep_co %>%
  left_join(unique_devrep, by = c("chemical1" = "chemical")) %>%
  rename(total_release1 = total_release)

devrep_co <- devrep_co %>%
  left_join(unique_devrep, by = c("chemical2" = "chemical")) %>%
  rename(total_release2 = total_release)

# Calculate percent corelease for heatmaps 
devrep_co <- devrep_co %>%
  mutate(perc_co = 100 * ((2 * unique_trifd_count) / (total_release1 + total_release2)))

# save data in csv
write.csv(devrep_co, file = "~/TSCA/CRA/CRA_Prioritization/Output/devrep_corelease.csv")

```


#co-releases in lkt
```{r}
# Get all unique chemicals
chemicals <- unique(lkt$chemical)

# Get all possible pairs of chemicals
chemical_pairs <- combn(chemicals, 2, simplify = FALSE)

# Loop through each pair and get the count of unique trifd
results <- lapply(chemical_pairs, function(pair) {
  count <- count_unique_trifd(lkt, pair[1], pair[2])
  data.frame(chemical1 = pair[1], chemical2 = pair[2], unique_trifd_count = count)
})

# Combine the results into a single dataframe
lkt_co <- do.call(rbind, results)

#make dataframe of number of facilities releasing each chemical
unique_lkt <- lkt %>%
  group_by(chemical) %>%
  summarise(total_release = n_distinct(trifd)) 

# Join dataframes to get total_release for chemical1 and chemical2
lkt_co <- lkt_co %>%
  left_join(unique_lkt, by = c("chemical1" = "chemical")) %>%
  rename(total_release1 = total_release)

lkt_co <- lkt_co %>%
  left_join(unique_lkt, by = c("chemical2" = "chemical")) %>%
  rename(total_release2 = total_release)

# Calculate percent corelease for heatmaps 
lkt_co <- lkt_co %>%
  mutate(perc_co = 100 * ((2 * unique_trifd_count) / (total_release1 + total_release2)))

# save data in csv
write.csv(lkt_co, file = "~/TSCA/CRA/CRA_Prioritization/Output/lkt_corelease.csv")

```



#carc heatmap 
```{r}
#remove unwanted columns from carc_co and reformat the dataframe
carc_co_wide <- carc_co %>%
  select (-unique_trifd_count, -total_release1, -total_release2) %>%
    mutate(
    chemical1 = ifelse(chemical1 == "4,4'-Methylenebis(2-chloroaniline)", "MBOCA", chemical1),
    chemical2 = ifelse(chemical2 == "4,4'-Methylenebis(2-chloroaniline)", "MBOCA", chemical2)
  ) %>%
  mutate(
    chemical1 = ifelse(chemical1 == "3,3'-Dichlorobenzidine dihydrochloride", "3,3'-Dichlorobenzidine 2HCl", chemical1),
    chemical2 = ifelse(chemical2 == "3,3'-Dichlorobenzidine dihydrochloride", "3,3'-Dichlorobenzidine 2HCl", chemical2)
  ) %>%
  pivot_wider(
    names_from = chemical1,    
    values_from = perc_co,   
    values_fill = list(unique_trifd_count = NA)  
  )

# Convert "chemical2" to row names
carc_co_wide <- as.data.frame(carc_co_wide)    
rownames(carc_co_wide) <- carc_co_wide$chemical2  
carc_co_wide$chemical2 <- NULL                 

#transpose columns and rows (output is a matrix, which is what is needed for pheatmap)
carc_co_transposed <- t(carc_co_wide)

#create breaks for colors
breaks <- c(0, 2, 4, 6, 8, 10, 20, 30, 50, 70)
colors <- brewer.pal(9, "Blues")

#make heatmap
pheatmap(carc_co_transposed, cluster_rows = FALSE, cluster_cols = FALSE, border_color = "grey15", breaks = breaks, display_numbers = FALSE, main = "Carcinogens", color = colors, na_col = "grey50", number_color = "grey15", cellheight=32,cellwidth=32, fontsize_number = 10, angle_col = 45, legend = T, filename = "~/TSCA/CRA/CRA_Prioritization/Output/carc_heat.png", width = 12, height = 9)
```


#cns heatmap 
```{r}
#remove unwanted columns from carc_co and reformat the dataframe
cns_co_wide <- cns_co %>%
  select (-unique_trifd_count, -total_release1, -total_release2) %>%
    mutate(
    chemical1 = ifelse(chemical1 == "Decabromodiphenyl oxide", "DecaBDE", chemical1),
    chemical2 = ifelse(chemical2 == "Decabromodiphenyl oxide", "DecaBDE", chemical2)
  ) %>%
  pivot_wider(
    names_from = chemical1,    
    values_from = perc_co,   
    values_fill = list(unique_trifd_count = NA)  
  )

# Convert "chemical2" to row names
cns_co_wide <- as.data.frame(cns_co_wide)    
rownames(cns_co_wide) <- cns_co_wide$chemical2  
cns_co_wide$chemical2 <- NULL                 

#transpose columns and rows (output is a matrix, which is what is needed for pheatmap)
cns_co_transposed <- t(cns_co_wide)

#create breaks for colors
breaks <- c(0, 2, 4, 6, 8, 10, 20, 30, 50, 70)
colors <- brewer.pal(9, "Blues")

#make heatmap
pheatmap(cns_co_transposed, cluster_rows = FALSE, cluster_cols = FALSE, border_color = "grey15", breaks = breaks, display_numbers = FALSE, main = "Central Nervous System Toxicants", color = colors, na_col = "grey50", number_color = "grey15", cellheight=32,cellwidth=32, fontsize_number = 10, angle_col = 45, legend = T, filename = "~/TSCA/CRA/CRA_Prioritization/Output/cns_heat.png", width = 10, height = 7)
```


#cardioresp heatmap 
```{r}
#remove unwanted columns from carc_co and reformat the dataframe
cardioresp_co_wide <- cardioresp_co %>%
  select (-unique_trifd_count, -total_release1, -total_release2) %>%
    mutate(
    chemical1 = ifelse(chemical1 == "Molybdenum trioxide", "Molybdenum", chemical1),
    chemical2 = ifelse(chemical2 == "Molybdenum trioxide", "Molybdenum", chemical2)) %>%
  pivot_wider(
    names_from = chemical1,    
    values_from = perc_co,   
    values_fill = list(unique_trifd_count = NA)  
  )

# Convert "chemical2" to row names
cardioresp_co_wide <- as.data.frame(cardioresp_co_wide)    
rownames(cardioresp_co_wide) <- cardioresp_co_wide$chemical2  
cardioresp_co_wide$chemical2 <- NULL         

#transpose columns and rows (output is a matrix, which is what is needed for pheatmap)
cardioresp_co_transposed <- t(cardioresp_co_wide)

#create breaks for colors
breaks <- c(0, 2, 4, 6, 8, 10, 20, 30, 50, 70)
colors <- brewer.pal(9, "Blues")

#make heatmap
pheatmap(cardioresp_co_transposed, cluster_rows = FALSE, cluster_cols = FALSE, border_color = "grey15", breaks = breaks, display_numbers = FALSE, main = "Cardiorespiratory Toxicants", color = colors, na_col = "grey50", number_color = "grey15", cellheight=32,cellwidth=32, fontsize_number = 10, angle_col = 45, legend = T, filename = "~/TSCA/CRA/CRA_Prioritization/Output/cardioresp_heat.png", width = 12, height = 9)
```


#devrep heatmap 
```{r}
#remove unwanted columns from carc_co and reformat the dataframe
devrep_co_wide <- devrep_co %>%
  select (-unique_trifd_count, -total_release1, -total_release2) %>%
    mutate(
    chemical1 = ifelse(chemical1 == "Molybdenum trioxide", "Molybdenum", chemical1),
    chemical2 = ifelse(chemical2 == "Molybdenum trioxide", "Molybdenum", chemical2)
  ) %>%
      mutate(
    chemical1 = ifelse(chemical1 == "4,4'-Isopropylidenediphenol", "Bisphenol A", chemical1),
    chemical2 = ifelse(chemical2 == "4,4'-Isopropylidenediphenol", "Bisphenol A", chemical2)
  ) %>%
    mutate(
    chemical1 = ifelse(chemical1 == "Decabromodiphenyl oxide", "DecaBDE", chemical1),
    chemical2 = ifelse(chemical2 == "Decabromodiphenyl oxide", "DecaBDE", chemical2)
  ) %>%
  pivot_wider(
    names_from = chemical1,    
    values_from = perc_co,   
    values_fill = list(unique_trifd_count = NA)  
  )

# Convert "chemical2" to row names
devrep_co_wide <- as.data.frame(devrep_co_wide)    
rownames(devrep_co_wide) <- devrep_co_wide$chemical2  
devrep_co_wide$chemical2 <- NULL                 

#transpose columns and rows (output is a matrix, which is what is needed for pheatmap)
devrep_co_transposed <- t(devrep_co_wide)

#create breaks for colors
breaks <- c(0, 2, 4, 6, 8, 10, 20, 30, 50, 70)
colors <- brewer.pal(9, "Blues")

#make heatmap
pheatmap(devrep_co_transposed, cluster_rows = FALSE, cluster_cols = FALSE, border_color = "grey15", breaks = breaks, display_numbers = FALSE, main = "Developmental/Reproductive Toxicants", color = colors, na_col = "grey50", number_color = "grey15", cellheight=32,cellwidth=32, fontsize_number = 10, angle_col = 45, legend = T, filename = "~/TSCA/CRA/CRA_Prioritization/Output/devrep_heat.png", width = 12, height = 9)
```


#lkt heatmap 
```{r}
#remove unwanted columns from carc_co and reformat the dataframe
lkt_co_wide <- lkt_co %>%
  select (-unique_trifd_count, -total_release1, -total_release2) %>%
    mutate(
    chemical1 = ifelse(chemical1 == "Molybdenum trioxide", "Molybdenum", chemical1),
    chemical2 = ifelse(chemical2 == "Molybdenum trioxide", "Molybdenum", chemical2)
  ) %>%
    mutate(
    chemical1 = ifelse(chemical1 == "Decabromodiphenyl oxide", "DecaBDE", chemical1),
    chemical2 = ifelse(chemical2 == "Decabromodiphenyl oxide", "DecaBDE", chemical2)
  ) %>%
      mutate(
    chemical1 = ifelse(chemical1 == "4,4'-Methylenebis(2-chloroaniline)", "MBOCA", chemical1),
    chemical2 = ifelse(chemical2 == "4,4'-Methylenebis(2-chloroaniline)", "MBOCA", chemical2)
  ) %>%
  pivot_wider(
    names_from = chemical1,    
    values_from = perc_co,   
    values_fill = list(unique_trifd_count = NA)  
  )

# Convert "chemical2" to row names
lkt_co_wide <- as.data.frame(lkt_co_wide)    
rownames(lkt_co_wide) <- lkt_co_wide$chemical2  
lkt_co_wide$chemical2 <- NULL                 

#transpose columns and rows (output is a matrix, which is what is needed for pheatmap)
lkt_co_transposed <- t(lkt_co_wide)

#create breaks for colors
breaks <- c(0, 2, 4, 6, 8, 10, 20, 30, 50, 70)
colors <- brewer.pal(9, "Blues")

#make heatmap
pheatmap(lkt_co_transposed, cluster_rows = FALSE, cluster_cols = FALSE, border_color = "grey15", breaks = breaks, display_numbers = FALSE, main = "Liver, Kidney, and/or Thyroid Toxicants", color = colors, na_col = "grey50", number_color = "grey15", cellheight=32,cellwidth=32, fontsize_number = 10, angle_col = 45, legend = T, filename = "~/TSCA/CRA/CRA_Prioritization/Output/lkt_heat.png", width = 13, height = 10)
```


#comparing lbs of chemicals released from single-chemical release facilities and multi-chemical release facilities
```{r}
# Identify unique and repeated trifd
tri_classified <- tri_endpoints %>%
  group_by(trifd) %>%
  mutate(trifd_count = n()) %>%  # Count occurrences of each trifd
  ungroup()

#Average for trifd that are unique (count == 1)
average_unique_trifd <- tri_classified %>%
  filter(trifd_count == 1) %>%
  group_by(chemical) %>%
  summarise(average_release_unique_trifd = mean(onsite_releases_2017_2022, na.rm = TRUE))

#Average for trifd that are repeated (count > 1)
average_repeated_trifd <- tri_classified %>%
  filter(trifd_count > 1) %>%
  group_by(chemical) %>%
  summarise(average_release_repeated_trifd = mean(onsite_releases_2017_2022, na.rm = TRUE))

#merge dataframes and save
releases_unique_repeated <- full_join(average_unique_trifd, average_repeated_trifd, by = "chemical")

write.csv(releases_unique_repeated, file = "~/TSCA/CRA/CRA_Prioritization/Output/unique.repeated.summary.csv")

#format data to run anova by classifying trifd as unique or repeated
tri_classified <- tri_endpoints %>%
  group_by(trifd) %>%
  mutate(trifd_count = n()) %>%  
  ungroup() %>%
  mutate(trifd_type = ifelse(trifd_count == 1, "unique", "repeated"))

#Run ANOVA
anova = aov(onsite_releases_2017_2022 ~ trifd_type*chemical, data = tri_classified)
summary(anova)
capture.output(anova, file = "~/TSCA/CRA/CRA_Prioritization/Output/release_anova")

tukey = TukeyHSD(x = anova, conf.level = 0.95)
capture.output(tukey, file = "~/TSCA/CRA/CRA_Prioritization/Output/release_tukey")

```

















