---
title: "Analysis of co-releases of chemicals in endpoint groups"
author: "Paige Varner"
date: "2024-09-13"
output: html_document
---


#load libraries
```{r}
library("dplyr")
library("tidyr")
library("bigrquery")
library("DBI")
library("pheatmap")
library("RColorBrewer")
library("Rmisc")
library("stats")
library("rstatix")
library("ggplot2")
library("car")
library("grid")
library("magick")
detach(package:Rmisc)
detach(package:plyr)

endpoints = read.csv("~/TSCA/CRA/CRA_Prioritization/Data/TSCA_endpoints.csv")

```


#query data from BQ (dataframe with TRI ID, facility name, chemicals, and total sum of on site releases for the 5 year span)
```{r}
	# set up database connection
	con <- dbConnect(
	  bigrquery::bigquery(),
	  project = "edf-aq-data",
	  dataset = "tri",
)

SQL = "	
	SELECT
	trifd,
	facility_name,
	chemical,
	SUM(on_site_release_total) as sum_onsite_releases_2017_2022
	FROM `edf-aq-data.tri.tri_basic_2017-2022` 
	GROUP BY trifd, facility_name, chemical"

tri = dbGetQuery(con,SQL)

tri = filter(tri, sum_onsite_releases_2017_2022 > 0)
```



#fixing multiple names for same chemical (need to add up releases for chemicals of the same name after fixing the names for individual trifds)
```{r}
#fix chemical names
tri$chemical <- gsub("Xylene \\(mixed isomers\\)", "Xylenes", tri$chemical)
tri$chemical = gsub("o-Xylene", "Xylenes", tri$chemical)
tri$chemical = gsub("p-Xylene", "Xylenes", tri$chemical)
tri$chemical = gsub("m-Xylene", "Xylenes", tri$chemical)

tri$chemical = gsub("Hydrogen cyanide", "Cyanide compounds", tri$chemical)

tri$chemical = gsub("Nickel compounds", "Nickel", tri$chemical)

tri$chemical = gsub("Arsenic compounds", "Arsenic", tri$chemical)

tri$chemical = gsub("Cadmium compounds", "Cadmium", tri$chemical)

tri$chemical = gsub("Chromium compounds", "Chromium", tri$chemical)

tri$chemical = gsub("Lead compounds", "Lead", tri$chemical)

tri$chemical = gsub("Cobalt compounds", "Cobalt", tri$chemical)

tri$chemical = gsub("Nonylphenol Ethoxylates", "Nonylphenol", tri$chemical)

tri$chemical = gsub("Barium compounds", "Barium", tri$chemical)


#add up chemicals with same name for each trifd to get total releases of those chemicals 
tri <- tri %>%
  group_by(trifd, chemical) %>%
  summarise(onsite_releases_2017_2022 = sum(sum_onsite_releases_2017_2022, na.rm = TRUE))

```

#add endpoint designations to tri data, filter based on endpoints
```{r}
tri_endpoints <- tri %>%
  left_join(endpoints, by = c("chemical" = "tri_name")) %>%
  filter(!is.na(analysis_name))

carc <- tri_endpoints %>%
  filter(carc %in% c("Known", "Probable"))

cns <- tri_endpoints %>%
  filter(cns %in% "yes")

cardioresp = tri_endpoints %>%
  filter(cardio == "yes" | resp == "yes")

devrep = tri_endpoints %>%
  filter(dev == "yes" | repro == "yes")

lkt = tri_endpoints %>%
  filter(lkt %in% "yes")
```


#calculate percent co-releases overall for each group: # of coreleases/unique facilities where # of coreleases is total number of facilities - unique facilities
```{r}
#cns = 52%
((length(cns[["trifd"]]) - length(unique(cns[["trifd"]]))) / length(unique(cns[["trifd"]])))*100

#carc = 42%
((length(carc[["trifd"]]) - length(unique(carc[["trifd"]]))) / length(unique(carc[["trifd"]])))*100

#cardioresp = 68%
((length(cardioresp[["trifd"]]) - length(unique(cardioresp[["trifd"]]))) / length(unique(cardioresp[["trifd"]])))*100

#lkt = 43%
((length(lkt[["trifd"]]) - length(unique(lkt[["trifd"]]))) / length(unique(lkt[["trifd"]])))*100

#devrep = 28%
((length(devrep[["trifd"]]) - length(unique(devrep[["trifd"]]))) / length(unique(devrep[["trifd"]])))*100
```



# Function to count unique trifd for a given pair of chemicals
```{r}
count_unique_trifd <- function(data, chem1, chem2) {
  data %>%
    filter(chemical %in% c(chem1, chem2)) %>%
    group_by(trifd) %>%
    filter(n_distinct(chemical) == 2) %>%  # Ensure both chemicals are present
    ungroup() %>%  # Ungroup to avoid errors in summarisation
    summarise(unique_trifd_count = n_distinct(trifd)) %>%  # Count unique trifd
    pull(unique_trifd_count)  # Extract the count as a numeric value
}
```


#co-releases in carc
```{r}
# Get all unique chemicals
chemicals <- unique(carc$chemical)

# Get all possible pairs of chemicals
chemical_pairs <- combn(chemicals, 2, simplify = FALSE)

# Loop through each pair and get the count of unique trifd
results <- lapply(chemical_pairs, function(pair) {
  count <- count_unique_trifd(carc, pair[1], pair[2])
  data.frame(chemical1 = pair[1], chemical2 = pair[2], unique_trifd_count = count)
})

# Combine the results into a single dataframe
carc_co <- do.call(rbind, results)

#make dataframe of number of facilities releasing each chemical
unique_carc <- carc %>%
  group_by(chemical) %>%
  summarise(total_release = n_distinct(trifd)) 

# Join dataframes to get total_release for chemical1 and chemical2
carc_co <- carc_co %>%
  left_join(unique_carc, by = c("chemical1" = "chemical")) %>%
  rename(total_release1 = total_release)

carc_co <- carc_co %>%
  left_join(unique_carc, by = c("chemical2" = "chemical")) %>%
  rename(total_release2 = total_release)

# Calculate percent corelease for heatmaps 
carc_co <- carc_co %>%
  mutate(perc_co = 100 * ((2 * unique_trifd_count) / (total_release1 + total_release2)))

# save data in csv
write.csv(carc_co, file = "~/TSCA/CRA/CRA_Prioritization/Output/carc_corelease.csv")
```


#co-releases in cns
```{r}
# Get all unique chemicals
chemicals <- unique(cns$chemical)

# Get all possible pairs of chemicals
chemical_pairs <- combn(chemicals, 2, simplify = FALSE)

# Loop through each pair and get the count of unique trifd
results <- lapply(chemical_pairs, function(pair) {
  count <- count_unique_trifd(cns, pair[1], pair[2])
  data.frame(chemical1 = pair[1], chemical2 = pair[2], unique_trifd_count = count)
})

# Combine the results into a single dataframe
cns_co <- do.call(rbind, results)

#make dataframe of number of facilities releasing each chemical
unique_cns <- cns %>%
  group_by(chemical) %>%
  summarise(total_release = n_distinct(trifd)) 

# Join dataframes to get total_release for chemical1 and chemical2
cns_co <- cns_co %>%
  left_join(unique_cns, by = c("chemical1" = "chemical")) %>%
  rename(total_release1 = total_release)

cns_co <- cns_co %>%
  left_join(unique_cns, by = c("chemical2" = "chemical")) %>%
  rename(total_release2 = total_release)

# Calculate percent corelease for heatmaps 
cns_co <- cns_co %>%
  mutate(perc_co = 100 * ((2 * unique_trifd_count) / (total_release1 + total_release2)))

# save data in csv
write.csv(cns_co, file = "~/TSCA/CRA/CRA_Prioritization/Output/cns_corelease.csv")

```


#co-releases in cardioresp
```{r}
# Get all unique chemicals
chemicals <- unique(cardioresp$chemical)

# Get all possible pairs of chemicals
chemical_pairs <- combn(chemicals, 2, simplify = FALSE)

# Loop through each pair and get the count of unique trifd
results <- lapply(chemical_pairs, function(pair) {
  count <- count_unique_trifd(cardioresp, pair[1], pair[2])
  data.frame(chemical1 = pair[1], chemical2 = pair[2], unique_trifd_count = count)
})

# Combine the results into a single dataframe
cardioresp_co <- do.call(rbind, results)

#make dataframe of number of facilities releasing each chemical
unique_cardioresp <- cardioresp %>%
  group_by(chemical) %>%
  summarise(total_release = n_distinct(trifd)) 

# Join dataframes to get total_release for chemical1 and chemical2
cardioresp_co <- cardioresp_co %>%
  left_join(unique_cardioresp, by = c("chemical1" = "chemical")) %>%
  rename(total_release1 = total_release)

cardioresp_co <- cardioresp_co %>%
  left_join(unique_cardioresp, by = c("chemical2" = "chemical")) %>%
  rename(total_release2 = total_release)

# Calculate percent corelease for heatmaps 
cardioresp_co <- cardioresp_co %>%
  mutate(perc_co = 100 * ((2 * unique_trifd_count) / (total_release1 + total_release2)))

# save data in csv
write.csv(cardioresp_co, file = "~/TSCA/CRA/CRA_Prioritization/Output/cardioresp_corelease.csv")

```


#co-releases in devrep
```{r}
# Get all unique chemicals
chemicals <- unique(devrep$chemical)

# Get all possible pairs of chemicals
chemical_pairs <- combn(chemicals, 2, simplify = FALSE)

# Loop through each pair and get the count of unique trifd
results <- lapply(chemical_pairs, function(pair) {
  count <- count_unique_trifd(devrep, pair[1], pair[2])
  data.frame(chemical1 = pair[1], chemical2 = pair[2], unique_trifd_count = count)
})

# Combine the results into a single dataframe
devrep_co <- do.call(rbind, results)

#make dataframe of number of facilities releasing each chemical
unique_devrep <- devrep %>%
  group_by(chemical) %>%
  summarise(total_release = n_distinct(trifd)) 

# Join dataframes to get total_release for chemical1 and chemical2
devrep_co <- devrep_co %>%
  left_join(unique_devrep, by = c("chemical1" = "chemical")) %>%
  rename(total_release1 = total_release)

devrep_co <- devrep_co %>%
  left_join(unique_devrep, by = c("chemical2" = "chemical")) %>%
  rename(total_release2 = total_release)

# Calculate percent corelease for heatmaps 
devrep_co <- devrep_co %>%
  mutate(perc_co = 100 * ((2 * unique_trifd_count) / (total_release1 + total_release2)))

# save data in csv
write.csv(devrep_co, file = "~/TSCA/CRA/CRA_Prioritization/Output/devrep_corelease.csv")

```


#co-releases in lkt
```{r}
# Get all unique chemicals
chemicals <- unique(lkt$chemical)

# Get all possible pairs of chemicals
chemical_pairs <- combn(chemicals, 2, simplify = FALSE)

# Loop through each pair and get the count of unique trifd
results <- lapply(chemical_pairs, function(pair) {
  count <- count_unique_trifd(lkt, pair[1], pair[2])
  data.frame(chemical1 = pair[1], chemical2 = pair[2], unique_trifd_count = count)
})

# Combine the results into a single dataframe
lkt_co <- do.call(rbind, results)

#make dataframe of number of facilities releasing each chemical
unique_lkt <- lkt %>%
  group_by(chemical) %>%
  summarise(total_release = n_distinct(trifd)) 

# Join dataframes to get total_release for chemical1 and chemical2
lkt_co <- lkt_co %>%
  left_join(unique_lkt, by = c("chemical1" = "chemical")) %>%
  rename(total_release1 = total_release)

lkt_co <- lkt_co %>%
  left_join(unique_lkt, by = c("chemical2" = "chemical")) %>%
  rename(total_release2 = total_release)

# Calculate percent corelease for heatmaps 
lkt_co <- lkt_co %>%
  mutate(perc_co = 100 * ((2 * unique_trifd_count) / (total_release1 + total_release2)))

# save data in csv
write.csv(lkt_co, file = "~/TSCA/CRA/CRA_Prioritization/Output/lkt_corelease.csv")

```



#carc heatmap 
```{r}
library(tidyverse)

# Prepare data in a wide format
carc_co_wide <- carc_co %>%
  select(-unique_trifd_count) %>%  # Keep total_release1 and total_release2
  mutate(
    chemical1 = ifelse(chemical1 == "4,4'-Methylenebis(2-chloroaniline)", "MBOCA", chemical1),
    chemical2 = ifelse(chemical2 == "4,4'-Methylenebis(2-chloroaniline)", "MBOCA", chemical2),
    chemical1 = ifelse(chemical1 == "3,3'-Dichlorobenzidine dihydrochloride", "3,3'-Dichlorobenzidine 2HCl", chemical1),
    chemical2 = ifelse(chemical2 == "3,3'-Dichlorobenzidine dihydrochloride", "3,3'-Dichlorobenzidine 2HCl", chemical2),
    # Create a version of chemical labels for processing, separate release values
    chemical1_label = paste0(chemical1, " (", total_release1, ")"),
    chemical2_label = paste0(chemical2, " (", total_release2, ")")
  ) %>%
  pivot_wider(
    names_from = chemical1_label,
    values_from = perc_co,
    values_fill = list(perc_co = NA)
  )

# Ensure chemical2 values are unique by appending a suffix if duplicates exist
carc_co_wide$chemical2_label <- make.unique(as.character(carc_co_wide$chemical2_label))

# Remove the suffixes from `chemical2_label` (e.g., "Lead (6824).1" -> "Lead (6824)")
carc_co_wide$chemical2_label <- sub("\\.\\d+$", "", carc_co_wide$chemical2_label)

# Remove total_release1 and total_release2 columns from the dataset before pivoting back to long format
carc_co_wide_clean <- carc_co_wide %>%
  select(-total_release1, -total_release2)

# Ensure that all columns (including chemical1_label) are treated as characters
carc_co_wide_clean[] <- lapply(carc_co_wide_clean, as.character)

# Convert back to long format without including total release columns
carc_co_long <- carc_co_wide_clean %>%
  pivot_longer(cols = -chemical2_label, names_to = "chemical1", values_to = "value") %>%
  drop_na() %>%
  rename(chemical2 = chemical2_label)  # Rename the column back to `chemical2`

# Order chemicals so they match the triangular shape
chemical_levels <- unique(c(carc_co_long$chemical1, carc_co_long$chemical2))
carc_co_long <- carc_co_long %>%
  mutate(
    chemical1 = factor(chemical1, levels = chemical_levels),
    chemical2 = factor(chemical2, levels = chemical_levels)
  )

# Keep only the lower triangle (remove upper diagonal)
carc_co_long <- carc_co_long %>%
  filter(as.numeric(chemical2) >= as.numeric(chemical1))

# Convert to numeric, non-numeric entries will become NA
carc_co_long$value <- as.numeric(carc_co_long$value)

# Check for NA values after conversion
sum(is.na(carc_co_long$value))

carc_co_long <- carc_co_long %>%
  drop_na(value)

# Create heatmap with black tile borders and updated labels
carc_heat = ggplot(carc_co_long, aes(x = chemical1, y = chemical2, fill = value)) +
  geom_tile(color = "gray", size = 0.5) +  # Black border around each tile
  scale_fill_gradientn(colors = brewer.pal(9, "Blues"), na.value = "transparent", name = "Co-Release Percentage") +  
  scale_x_discrete(position = "top") +  # Moves x-axis labels to the top
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 0, vjust = 0, size = 10, color = "black"),  # Set x-axis text color to black
    axis.text.y = element_text(size = 10, color = "black"),  # Set y-axis text color to black
    axis.title.x = element_blank(),  
    axis.title.y = element_blank(),  
    panel.grid = element_blank()  
  )


carc_heat

# Save the heatmap
ggsave(filename = "carc_heat.png", device = "png", plot = carc_heat, path = "~/TSCA/CRA/CRA_Prioritization/Output/", width = 7.5, height = 4.5)


```


#cns heatmap 
```{r}
# Prepare data in a wide format
cns_co_wide <- cns_co %>%
  select(-unique_trifd_count) %>%  # Keep total_release1 and total_release2
  mutate(
    chemical1 = ifelse(chemical1 == "Decabromodiphenyl oxide", "DecaBDE", chemical1),
    chemical2 = ifelse(chemical2 == "Decabromodiphenyl oxide", "DecaBDE", chemical2),
    chemical1 = ifelse(chemical1 == "Cyanide compounds", "Cyanide", chemical1),
    chemical2 = ifelse(chemical2 == "Cyanide compounds", "Cyanide", chemical2),
    # Create a version of chemical labels for processing, separate release values
    chemical1_label = paste0(chemical1, " (", total_release1, ")"),
    chemical2_label = paste0(chemical2, " (", total_release2, ")")
  ) %>%
  pivot_wider(
    names_from = chemical1_label,
    values_from = perc_co,
    values_fill = list(perc_co = NA)
  )

# Ensure chemical2 values are unique by appending a suffix if duplicates exist
cns_co_wide$chemical2_label <- make.unique(as.character(cns_co_wide$chemical2_label))

# Remove the suffixes from `chemical2_label` (e.g., "Lead (6824).1" -> "Lead (6824)")
cns_co_wide$chemical2_label <- sub("\\.\\d+$", "", cns_co_wide$chemical2_label)

# Remove total_release1 and total_release2 columns from the dataset before pivoting back to long format
cns_co_wide_clean <- cns_co_wide %>%
  select(-total_release1, -total_release2)

# Ensure that all columns (including chemical1_label) are treated as characters
cns_co_wide_clean[] <- lapply(cns_co_wide_clean, as.character)

# Convert back to long format without including total release columns
cns_co_long <- cns_co_wide_clean %>%
  pivot_longer(cols = -chemical2_label, names_to = "chemical1", values_to = "value") %>%
  drop_na() %>%
  rename(chemical2 = chemical2_label)  # Rename the column back to `chemical2`

# Order chemicals so they match the triangular shape
chemical_levels <- unique(c(cns_co_long$chemical1, cns_co_long$chemical2))
cns_co_long <- cns_co_long %>%
  mutate(
    chemical1 = factor(chemical1, levels = chemical_levels),
    chemical2 = factor(chemical2, levels = chemical_levels)
  )

# Keep only the lower triangle (remove upper diagonal)
cns_co_long <- cns_co_long %>%
  filter(as.numeric(chemical2) >= as.numeric(chemical1))

# Convert to numeric, non-numeric entries will become NA
cns_co_long$value <- as.numeric(cns_co_long$value)

# Check for NA values after conversion
sum(is.na(cns_co_long$value))

cns_co_long <- cns_co_long %>%
  drop_na(value)

# Create heatmap with black tile borders and updated labels
cns_heat = ggplot(cns_co_long, aes(x = chemical1, y = chemical2, fill = value)) +
  geom_tile(color = "gray", size = 0.5) +  # Black border around each tile
  scale_fill_gradientn(colors = brewer.pal(9, "Blues"), na.value = "transparent", name = "Co-Release Percentage") +  
  scale_x_discrete(position = "top") +  # Moves x-axis labels to the top
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 0, vjust = 0, size = 10, color = "black"),  # Set x-axis text color to black
    axis.text.y = element_text(size = 10, color = "black"),  # Set y-axis text color to black
    axis.title.x = element_blank(),  
    axis.title.y = element_blank(),  
    panel.grid = element_blank()  
  )

cns_heat

# Save the heatmap
ggsave(filename = "cns_heat.png", device = "png", plot = cns_heat, path = "~/TSCA/CRA/CRA_Prioritization/Output/", width = 8, height = 5)

```


#cardioresp heatmap 
```{r}
# Prepare data in a wide format
cardioresp_co_wide <- cardioresp_co %>%
  select(-unique_trifd_count) %>%  # Keep total_release1 and total_release2
  mutate(
    chemical1 = ifelse(chemical1 == "Molybdenum trioxide", "Molybdenum", chemical1),
    chemical2 = ifelse(chemical2 == "Molybdenum trioxide", "Molybdenum", chemical2),
    # Create a version of chemical labels for processing, separate release values
    chemical1_label = paste0(chemical1, " (", total_release1, ")"),
    chemical2_label = paste0(chemical2, " (", total_release2, ")")
  ) %>%
  pivot_wider(
    names_from = chemical1_label,
    values_from = perc_co,
    values_fill = list(perc_co = NA)
  )

# Ensure chemical2 values are unique by appending a suffix if duplicates exist
cardioresp_co_wide$chemical2_label <- make.unique(as.character(cardioresp_co_wide$chemical2_label))

# Remove the suffixes from `chemical2_label` (e.g., "Lead (6824).1" -> "Lead (6824)")
cardioresp_co_wide$chemical2_label <- sub("\\.\\d+$", "", cardioresp_co_wide$chemical2_label)

# Remove total_release1 and total_release2 columns from the dataset before pivoting back to long format
cardioresp_co_wide_clean <- cardioresp_co_wide %>%
  select(-total_release1, -total_release2)

# Ensure that all columns (including chemical1_label) are treated as characters
cardioresp_co_wide_clean[] <- lapply(cardioresp_co_wide_clean, as.character)

# Convert back to long format without including total release columns
cardioresp_co_long <- cardioresp_co_wide_clean %>%
  pivot_longer(cols = -chemical2_label, names_to = "chemical1", values_to = "value") %>%
  drop_na() %>%
  rename(chemical2 = chemical2_label)  # Rename the column back to `chemical2`

# Order chemicals so they match the triangular shape
chemical_levels <- unique(c(cardioresp_co_long$chemical1, cardioresp_co_long$chemical2))
cardioresp_co_long <- cardioresp_co_long %>%
  mutate(
    chemical1 = factor(chemical1, levels = chemical_levels),
    chemical2 = factor(chemical2, levels = chemical_levels)
  )

# Keep only the lower triangle (remove upper diagonal)
cardioresp_co_long <- cardioresp_co_long %>%
  filter(as.numeric(chemical2) >= as.numeric(chemical1))

# Convert to numeric, non-numeric entries will become NA
cardioresp_co_long$value <- as.numeric(cardioresp_co_long$value)

# Check for NA values after conversion
sum(is.na(cardioresp_co_long$value))

cardioresp_co_long <- cardioresp_co_long %>%
  drop_na(value)

# Create heatmap with black tile borders and updated labels
cardioresp_heat = ggplot(cardioresp_co_long, aes(x = chemical1, y = chemical2, fill = value)) +
  geom_tile(color = "gray", size = 0.5) +  # Black border around each tile
  scale_fill_gradientn(colors = brewer.pal(9, "Blues"), na.value = "transparent", name = "Co-Release Percentage") +  
  scale_x_discrete(position = "top") +  # Moves x-axis labels to the top
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 0, vjust = 0, size = 10, color = "black"),  # Set x-axis text color to black
    axis.text.y = element_text(size = 10, color = "black"),  # Set y-axis text color to black
    axis.title.x = element_blank(),  
    axis.title.y = element_blank(),  
    panel.grid = element_blank()  
  )

cardioresp_heat

# Save the heatmap
ggsave(filename = "cardioresp_heat.png", device = "png", plot = cardioresp_heat, path = "~/TSCA/CRA/CRA_Prioritization/Output/", width = 8, height = 5)

```


#devrep heatmap 
```{r}

# Prepare data in a wide format
devrep_co_wide <- devrep_co %>%
  select(-unique_trifd_count) %>%  # Keep total_release1 and total_release2
  mutate(
    chemical1 = ifelse(chemical1 == "Molybdenum trioxide", "Molybdenum", chemical1),
    chemical2 = ifelse(chemical2 == "Molybdenum trioxide", "Molybdenum", chemical2),
    chemical1 = ifelse(chemical1 == "4,4'-Isopropylidenediphenol", "Bisphenol A", chemical1),
    chemical2 = ifelse(chemical2 == "4,4'-Isopropylidenediphenol", "Bisphenol A", chemical2),
    chemical1 = ifelse(chemical1 == "Decabromodiphenyl oxide", "DecaBDE", chemical1),
    chemical2 = ifelse(chemical2 == "Decabromodiphenyl oxide", "DecaBDE", chemical2),
    chemical1 = ifelse(chemical1 == "Hexachloro-1,3-butadiene", "Hexachlorobutadiene", chemical1),
    chemical2 = ifelse(chemical2 == "Hexachloro-1,3-butadiene", "Hexachlorobutadiene", chemical2),
    # Create a version of chemical labels for processing, separate release values
    chemical1_label = paste0(chemical1, " (", total_release1, ")"),
    chemical2_label = paste0(chemical2, " (", total_release2, ")")
  ) %>%
  pivot_wider(
    names_from = chemical1_label,
    values_from = perc_co,
    values_fill = list(perc_co = NA)
  )

# Ensure chemical2 values are unique by appending a suffix if duplicates exist
devrep_co_wide$chemical2_label <- make.unique(as.character(devrep_co_wide$chemical2_label))

# Remove the suffixes from `chemical2_label` (e.g., "Lead (6824).1" -> "Lead (6824)")
devrep_co_wide$chemical2_label <- sub("\\.\\d+$", "", devrep_co_wide$chemical2_label)

# Remove total_release1 and total_release2 columns from the dataset before pivoting back to long format
devrep_co_wide_clean <- devrep_co_wide %>%
  select(-total_release1, -total_release2)

# Ensure that all columns (including chemical1_label) are treated as characters
devrep_co_wide_clean[] <- lapply(devrep_co_wide_clean, as.character)

# Convert back to long format without including total release columns
devrep_co_long <- devrep_co_wide_clean %>%
  pivot_longer(cols = -chemical2_label, names_to = "chemical1", values_to = "value") %>%
  drop_na() %>%
  rename(chemical2 = chemical2_label)  # Rename the column back to `chemical2`

# Order chemicals so they match the triangular shape
chemical_levels <- unique(c(devrep_co_long$chemical1, devrep_co_long$chemical2))
devrep_co_long <- devrep_co_long %>%
  mutate(
    chemical1 = factor(chemical1, levels = chemical_levels),
    chemical2 = factor(chemical2, levels = chemical_levels)
  )

# Keep only the lower triangle (remove upper diagonal)
devrep_co_long <- devrep_co_long %>%
  filter(as.numeric(chemical2) >= as.numeric(chemical1))

# Convert to numeric, non-numeric entries will become NA
devrep_co_long$value <- as.numeric(devrep_co_long$value)

# Check for NA values after conversion
sum(is.na(devrep_co_long$value))

devrep_co_long <- devrep_co_long %>%
  drop_na(value)

# Create heatmap with black tile borders and updated labels
devrep_heat = ggplot(devrep_co_long, aes(x = chemical1, y = chemical2, fill = value)) +
  geom_tile(color = "gray", size = 0.5) +  # Black border around each tile
  scale_fill_gradientn(colors = brewer.pal(9, "Blues"), na.value = "transparent", name = "Co-Release Percentage") +  
  scale_x_discrete(position = "top") +  # Moves x-axis labels to the top
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 0, vjust = 0, size = 10, color = "black"),  # Set x-axis text color to black
    axis.text.y = element_text(size = 10, color = "black"),  # Set y-axis text color to black
    axis.title.x = element_blank(),  
    axis.title.y = element_blank(),  
    panel.grid = element_blank()  
  )

devrep_heat

# Save the heatmap
ggsave(filename = "devrep_heat.png", device = "png", plot = devrep_heat, path = "~/TSCA/CRA/CRA_Prioritization/Output/", width = 8, height = 5)


```


#lkt heatmap 
```{r}
  
# Prepare data in a wide format
lkt_co_wide <- lkt_co %>%
  select(-unique_trifd_count) %>%  # Keep total_release1 and total_release2
  mutate(
    chemical1 = ifelse(chemical1 == "Molybdenum trioxide", "Molybdenum", chemical1),
    chemical2 = ifelse(chemical2 == "Molybdenum trioxide", "Molybdenum", chemical2),
    chemical1 = ifelse(chemical1 == "Decabromodiphenyl oxide", "DecaBDE", chemical1),
    chemical2 = ifelse(chemical2 == "Decabromodiphenyl oxide", "DecaBDE", chemical2),
    chemical1 = ifelse(chemical1 == "4,4'-Methylenebis(2-chloroaniline)", "MBOCA", chemical1),
    chemical2 = ifelse(chemical2 == "4,4'-Methylenebis(2-chloroaniline)", "MBOCA", chemical2),
      chemical1 = ifelse(chemical1 == "Hexachloro-1,3-butadiene", "Hexachlorobutadiene", chemical1),
    chemical2 = ifelse(chemical2 == "Hexachloro-1,3-butadiene", "Hexachlorobutadiene", chemical2),
    # Create a version of chemical labels for processing, separate release values
    chemical1_label = paste0(chemical1, " (", total_release1, ")"),
    chemical2_label = paste0(chemical2, " (", total_release2, ")")
  ) %>%
  pivot_wider(
    names_from = chemical1_label,
    values_from = perc_co,
    values_fill = list(perc_co = NA)
  )

# Ensure chemical2 values are unique by appending a suffix if duplicates exist
lkt_co_wide$chemical2_label <- make.unique(as.character(lkt_co_wide$chemical2_label))

# Remove the suffixes from `chemical2_label` (e.g., "Lead (6824).1" -> "Lead (6824)")
lkt_co_wide$chemical2_label <- sub("\\.\\d+$", "", lkt_co_wide$chemical2_label)

# Remove total_release1 and total_release2 columns from the dataset before pivoting back to long format
lkt_co_wide_clean <- lkt_co_wide %>%
  select(-total_release1, -total_release2)

# Ensure that all columns (including chemical1_label) are treated as characters
lkt_co_wide_clean[] <- lapply(lkt_co_wide_clean, as.character)

# Convert back to long format without including total release columns
lkt_co_long <- lkt_co_wide_clean %>%
  pivot_longer(cols = -chemical2_label, names_to = "chemical1", values_to = "value") %>%
  drop_na() %>%
  rename(chemical2 = chemical2_label)  # Rename the column back to `chemical2`

# Order chemicals so they match the triangular shape
chemical_levels <- unique(c(lkt_co_long$chemical1, lkt_co_long$chemical2))
lkt_co_long <- lkt_co_long %>%
  mutate(
    chemical1 = factor(chemical1, levels = chemical_levels),
    chemical2 = factor(chemical2, levels = chemical_levels)
  )

# Keep only the lower triangle (remove upper diagonal)
lkt_co_long <- lkt_co_long %>%
  filter(as.numeric(chemical2) >= as.numeric(chemical1))

# Convert to numeric, non-numeric entries will become NA
lkt_co_long$value <- as.numeric(lkt_co_long$value)

# Check for NA values after conversion
sum(is.na(lkt_co_long$value))

lkt_co_long <- lkt_co_long %>%
  drop_na(value)

# Create heatmap with black tile borders and updated labels
lkt_heat = ggplot(lkt_co_long, aes(x = chemical1, y = chemical2, fill = value)) +
  geom_tile(color = "gray", size = 0.5) +  # Black border around each tile
  scale_fill_gradientn(colors = brewer.pal(9, "Blues"), na.value = "transparent", name = "Co-Release Percentage") +  
  scale_x_discrete(position = "top") +  # Moves x-axis labels to the top
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 0, vjust = 0, size = 10, color = "black"),  # Set x-axis text color to black
    axis.text.y = element_text(size = 10, color = "black"),  # Set y-axis text color to black
    axis.title.x = element_blank(),  
    axis.title.y = element_blank(),  
    panel.grid = element_blank()  
  )

lkt_heat

# Save the heatmap
ggsave(filename = "lkt_heat.png", device = "png", plot = lkt_heat, path = "~/TSCA/CRA/CRA_Prioritization/Output/", width = 8, height = 5)



```


#comparing lbs of chemicals released from single-chemical release facilities and multi-chemical release facilities
```{r}
# Identify unique and repeated trifd
tri_classified <- tri_endpoints %>%
  group_by(trifd) %>%
  mutate(trifd_count = n()) %>%  # Count occurrences of each trifd
  ungroup()

# Average and sample size for trifd that are unique (count == 1)
average_unique_trifd <- tri_classified %>%
  filter(trifd_count == 1) %>%
  group_by(chemical) %>%
  summarise(
    average_release_unique_trifd = mean(onsite_releases_2017_2022, na.rm = TRUE),
    sample_size_unique = n()
  )

# Average and sample size for trifd that are repeated (count > 1)
average_repeated_trifd <- tri_classified %>%
  filter(trifd_count > 1) %>%
  group_by(chemical) %>%
  summarise(
    average_release_repeated_trifd = mean(onsite_releases_2017_2022, na.rm = TRUE),
    sample_size_repeated = n()
  )

# Merge dataframes and save
releases_unique_repeated <- full_join(average_unique_trifd, average_repeated_trifd, by = "chemical")

write.csv(releases_unique_repeated, file = "~/TSCA/CRA/CRA_Prioritization/Output/unique.repeated.summary.csv")


#format data to run anova by classifying trifd as unique or repeated
tri_classified <- tri_endpoints %>%
  group_by(trifd) %>%
  mutate(trifd_count = n()) %>%  
  ungroup() %>%
  mutate(trifd_type = ifelse(trifd_count == 1, "unique", "repeated"))


#Run ANOVA
options(max.print=10000000)

anova = aov(onsite_releases_2017_2022 ~ trifd_type*chemical, data = tri_classified)
summary(anova)
capture.output(anova, file = "~/TSCA/CRA/CRA_Prioritization/Output/release_anova")

#Shapiro-Wilk test for normality (p < 0.05, data is not normally distributed); this data is extreeeemely skewed and needs to be log-normalized
shapiro.test(sample(anova$residuals, 5000))

hist(tri_classified$onsite_releases_2017_2022)

#Levene's Test for homogeneity of variances (p > 0.05, variances are homogenous)
leveneTest(onsite_releases_2017_2022 ~ trifd_type*chemical, data = tri_classified)

#log-transform and check distribution with histogram and tests (much more normal, at least visually)
tri_classified$log_onsite_releases = log(tri_classified$onsite_releases_2017_2022)
hist(tri_classified$log_onsite_releases)
leveneTest(log_onsite_releases ~ trifd_type*chemical, data = tri_classified)

#run ANOVA with transformed values - now everything is very significant
anova_log = aov(log_onsite_releases ~ trifd_type*chemical, data = tri_classified)
summary(anova_log)
capture.output(anova_log, file = "~/TSCA/CRA/CRA_Prioritization/Output/release_anova_transformed")

#posthoc tukey
tukey = TukeyHSD(x = anova_log, conf.level = 0.95)
capture.output(tukey, file = "~/TSCA/CRA/CRA_Prioritization/Output/release_tukey_transformed")

#overall averages
repeated = tri_classified %>%
  filter(trifd_type == "repeated")
mean(repeated$onsite_releases_2017_2022) #319,809.4

unique = tri_classified %>%
  filter(trifd_type == "unique")
mean(unique$onsite_releases_2017_2022) #33,101.53
```



#combine images for MS
```{r}
# Set working directory
setwd("~/TSCA/CRA/CRA_Prioritization/Output/")

# Define file names and labels
files <- c("carc_heat.png", "cns_heat.png", "cardioresp_heat.png", "devrep_heat.png", "lkt_heat.png")
labels <- c("A", "B", "C", "D", "E")

# Get max width to standardize all images
max_width <- max(sapply(files, function(f) image_info(image_read(f))$width))

# Read, flatten, resize, annotate
imgs <- mapply(function(file, label) {
  img <- image_read(file)
  img <- image_background(img, "white", flatten = TRUE)
  img <- image_resize(img, paste0(max_width))
  image_annotate(img, label, size = 50, gravity = "northwest", color = "black", location = "+20+20")
}, files, labels, SIMPLIFY = FALSE)

# Combine images vertically
combined <- image_append(image_join(imgs), stack = TRUE)

# Ensure white background
combined <- image_background(combined, "white", flatten = TRUE)

# Save as high-resolution PDF (300 DPI)
image_write(combined, path = "combined_heat.pdf", density = "300x300", format = "pdf")


```








