---
title: "CVI"
author: "Paige Varner"
date: "2023-09-21"
output: html_document
---

#load libraries
```{r}
library("dplyr")
library("bigrquery")
library("DBI")
library("pheatmap")
library("RColorBrewer")
library("Rmisc")
library("stats")
library("rstatix")
library("ggplot2")
library("tigris")
library("sf")
library("tidycensus")
```


#query data from BQ (dataframe with TRI ID, facility name, chemicals, lat/long, and total sum of on site releases for the 5 year span)
```{r}
	# set up database connection
	con <- dbConnect(
	  bigrquery::bigquery(),
	  project = "edf-aq-data",
	  dataset = "tri",
)

SQL = "	
	SELECT
	trifd,
	facility_name,
	chemical,
	latitude,
	longitude,
	SUM(on_site_release_total) as sum_onsite_releases
	FROM `edf-aq-data.tri.tri_basic_2016-2021` 
	GROUP BY trifd, facility_name, chemical,latitude,longitude"

tri = dbGetQuery(con,SQL)

tri = filter(tri, sum_onsite_releases > 0)
```


#get census tract GEOIDs for all TRI facilities based on long/lat
```{r}
cvi = read.csv("~/TSCA/CRA/CRA_Prioritization/Data/cvi_scores.csv")
colnames(cvi)[1] = "GEOID"

#getting census tract info for all TRI facilities
tracts = tracts(state = NULL, cb = TRUE)

tri.points <- tri %>%
  filter(!is.na(latitude)) %>%  # remove missing
  st_as_sf(coords=c('longitude', 'latitude'), crs=st_crs(tracts))

tri.tracts = st_join(tri.points, tracts)
```


#join CVI data to TRI data based on GEOID
```{r}
tri_cvi = merge(tri.tracts, cvi, by = "GEOID")
```


#fix multiple chemical names
```{r}
tri_cvi$chemical = gsub("Hydrogen cyanide", "Cyanide compounds", tri_cvi$chemical)
tri_cvi$chemical = gsub("m-Xylene", "Xylene (mixed isomers)", tri_cvi$chemical)
tri_cvi$chemical = gsub("o-Xylene", "Xylene (mixed isomers)", tri_cvi$chemical)
tri_cvi$chemical = gsub("p-Xylene", "Xylene (mixed isomers)", tri_cvi$chemical)
tri_cvi$chemical = gsub("Cadmium compounds", "Cadmium", tri_cvi$chemical)
tri_cvi$chemical = gsub("Nickel compounds", "Nickel", tri_cvi$chemical)
tri_cvi$chemical = gsub("Arsenic compounds", "Arsenic", tri_cvi$chemical)
tri_cvi$chemical = gsub("Nickel compounds", "Nickel", tri_cvi$chemical)
tri_cvi$chemical = gsub("Lead compounds", "Lead", tri_cvi$chemical)
tri_cvi$chemical = gsub("Barium compounds", "Barium", tri_cvi$chemical)
tri_cvi$chemical = gsub("Cobalt compounds", "Cobalt", tri_cvi$chemical)
tri_cvi$chemical = gsub("Nonylphenol Ethoxylates", "Nonylphenol", tri_cvi$chemical)


tri_cvi = tri_cvi[!duplicated(tri_cvi[c("trifd", "chemical")]),]

write.csv(tri_cvi, "~/TSCA/CRA/CRA_Prioritization/Data/tri_cvi.csv")
```



#Get average baseline CVI score for each chemical and endpoint group
```{r}
baseline_bychem = summarySE(tri_cvi, groupvars = "chemical", measurevar = "Baseline.Vulnerabilities")

#baseline for CNS chemicals
baseline_CNS = filter(baseline_bychem, chemical == "Cyanide compounds" | chemical == "Hydrogen cyanide" |  chemical == "Ethylbenzene" | chemical == "m-Xylene" | chemical == "o-Xylene" | chemical == "p-Xylene" | chemical == "Xylene (mixed isomers)")

#baseline for Carcinogens with CNS
baseline_carc_CNS = filter(baseline_bychem, chemical == "Acetaldehyde" | chemical == "Acrylonitrile"|chemical == "Aniline"|chemical == "Styrene"|chemical == "Benzene"|chemical == "Bromoform"|chemical == "Vinyl chloride")

#baseline for Carcinogens
baseline_carc = filter(baseline_bychem, chemical == "Creosote"|chemical == "3,3'-Dichlorobenzidine"|chemical == "3,3'-Dichlorobenzidine dihydrochloride"|chemical == "4,4'-Methylenebis(2-chloroaniline)"|chemical == "N-Nitrosodiphenylamine")

#baseline for Carcinogenetic metals
baseline_carc_metals = filter(baseline_bychem, chemical == "Cadmium" | chemical == "Cadmium compounds"|chemical == "Nickel" | chemical == "Nickel compounds"|chemical == "Arsenic" | chemical == "Arsenic compounds"|chemical == "Chromium" | chemical == "Chromium compounds"|chemical == "Lead" | chemical == "Lead compounds")

#baseline for cardiorespiratory
baseline_cardio = filter(baseline_bychem, chemical == "Barium" | chemical == "Barium compounds"|chemical == "Cobalt" | chemical == "Cobalt compounds"|chemical == "Molybdenum trioxide"|chemical == "Naphthalene"|chemical == "Maleic anhydride"|chemical == "1-Hexadecanol"|chemical == "Hydrogen fluoride") 

#baseline for liver, kidney, thyroid
baseline_lkt = filter(baseline_bychem, chemical == "Hexachloro-1,3-butadiene")

#baseline for edcs
baseline_edcs = filter(baseline_bychem, chemical == "Nonylphenol" | chemical == "Nonylphenol Ethoxylates"|chemical == "4,4'-Isopropylidenediphenol"|chemical == "Decabromodiphenyl oxide")

```

#add groupings to baseline vulnerability dfs so I can add back into one big df
```{r}
baseline_carc$endpoint = "carc"
baseline_carc_CNS$endpoint = "carc_CNS"
baseline_carc_metals$endpoint = "carc_metals"
baseline_cardio$endpoint = "cardio"
baseline_CNS$endpoint = "CNS"
baseline_edcs$endpoint = "edcs"
baseline_lkt$endpoint = "lkt"

all_baseline = rbind(baseline_carc, baseline_carc_CNS, baseline_carc_metals, baseline_cardio, baseline_CNS, baseline_edcs, baseline_lkt)

tri_cvi_baseline = merge(tri_cvi, all_baseline[,c("chemical", "endpoint")], by = "chemical")
```


#summary and stats to compare between end groups and graph
```{r}
tri_cvi_aov_baseline = aov(Baseline.Vulnerabilities ~ endpoint, data = tri_cvi_baseline)
tri_cvi_aov_baseline
tri_cvi_tukey_baseline = TukeyHSD(x = tri_cvi_aov_baseline, conf.level = 0.95)
tri_cvi_tukey_baseline
capture.output(tri_cvi_tukey_baseline, file = "~/TSCA/CRA/CRA_Prioritization/Output/tri_cvi_tukey")

#Boxplot 
endpoints = c("Carcinogens", "Carcinogenic CNS", "Carcinogenetic Metals", "Cardiorespiratory", "CNS", "EDCs", "Liver, Kidney, Thyroid")

tri_cvi_box_baseline = ggplot(tri_cvi_baseline, aes(x = endpoint, y = Baseline.Vulnerabilities)) + geom_boxplot(position = position_dodge()) + theme_bw() + ylab("Baseline Vulnerability") + xlab("Endpoint Group") + scale_x_discrete(labels = endpoints) + theme(axis.text.x = element_text(size = 10, angle = 45, hjust = 1))

tri_cvi_box_baseline

ggsave(filename = "tri_cvi_box_baseline.png", device = "png", plot = tri_cvi_box_baseline, path = "~/TSCA/CRA/CRA_Prioritization/Output/", width = 10, height = 5)

```


#filter tri_cvi_baseline for each endpoint group (without summary numbers, like above in baseline_endpoint dfs)
```{r}
cvi_CNS = filter(tri_cvi, chemical == "Cyanide compounds" | chemical == "Hydrogen cyanide" |  chemical == "Ethylbenzene" | chemical == "m-Xylene" | chemical == "o-Xylene" | chemical == "p-Xylene" | chemical == "Xylene (mixed isomers)")

cvi_carc_CNS = filter(tri_cvi, chemical == "Acetaldehyde" | chemical == "Acrylonitrile"|chemical == "Aniline"|chemical == "Styrene"|chemical == "Benzene"|chemical == "Bromoform"|chemical == "Vinyl chloride")

cvi_carc = filter(tri_cvi, chemical == "Creosote"|chemical == "3,3'-Dichlorobenzidine"|chemical == "3,3'-Dichlorobenzidine dihydrochloride"|chemical == "4,4'-Methylenebis(2-chloroaniline)"|chemical == "N-Nitrosodiphenylamine")

cvi_carc_metals = filter(tri_cvi, chemical == "Cadmium" | chemical == "Cadmium compounds"|chemical == "Nickel" | chemical == "Nickel compounds"|chemical == "Arsenic" | chemical == "Arsenic compounds"|chemical == "Chromium" | chemical == "Chromium compounds"|chemical == "Lead" | chemical == "Lead compounds")

cvi_cardio = filter(tri_cvi, chemical == "Barium" | chemical == "Barium compounds"|chemical == "Cobalt" | chemical == "Cobalt compounds"|chemical == "Molybdenum trioxide"|chemical == "Naphthalene"|chemical == "Maleic anhydride"|chemical == "1-Hexadecanol"|chemical == "Hydrogen fluoride") 

cvi_lkt = filter(tri_cvi, chemical == "Hexachloro-1,3-butadiene")

cvi_edcs = filter(tri_cvi, chemical == "Nonylphenol" | chemical == "Nonylphenol Ethoxylates"|chemical == "4,4'-Isopropylidenediphenol"|chemical == "Decabromodiphenyl oxide")
```


#boxplots for baseline vulnerability of each chemical within endpoint groups
```{r}
#carcinogens
carc_box_baseline = ggplot(cvi_carc, aes(x = chemical, y = Baseline.Vulnerabilities)) + geom_boxplot(position = position_dodge()) + theme_bw() + ylab("Baseline Vulnerability") + xlab("Chemical")  + theme(axis.text.x = element_text(size = 10, angle = 45, hjust = 1)) + scale_y_continuous(breaks = seq(0.2, 0.8, 0.1), limits = c(0.2,0.8))

carc_box_baseline

ggsave(filename = "carc_box_baseline.png", device = "png", plot = carc_box_baseline, path = "~/TSCA/CRA/CRA_Prioritization/Output/", width = 10, height = 5)

#CNS
CNS_box_baseline = ggplot(cvi_CNS, aes(x = chemical, y = Baseline.Vulnerabilities)) + geom_boxplot(position = position_dodge()) + theme_bw() + ylab("Baseline Vulnerability") + xlab("Chemical")  + theme(axis.text.x = element_text(size = 10, angle = 45, hjust = 1)) + scale_y_continuous(breaks = seq(0.2, 0.8, 0.1), limits = c(0.2,0.8))

CNS_box_baseline

ggsave(filename = "CNS_box_baseline.png", device = "png", plot = CNS_box_baseline, path = "~/TSCA/CRA/CRA_Prioritization/Output/", width = 10, height = 5)

#carcinogenic metals
carc_metals_box_baseline = ggplot(cvi_carc_metals, aes(x = chemical, y = Baseline.Vulnerabilities)) + geom_boxplot(position = position_dodge()) + theme_bw() + ylab("Baseline Vulnerability") + xlab("Chemical")  + theme(axis.text.x = element_text(size = 10, angle = 45, hjust = 1)) + scale_y_continuous(breaks = seq(0.2, 0.8, 0.1), limits = c(0.2,0.8))

carc_metals_box_baseline

ggsave(filename = "carc_metals_box_baseline.png", device = "png", plot = carc_metals_box_baseline, path = "~/TSCA/CRA/CRA_Prioritization/Output/", width = 10, height = 5)

#carcinogenic with CNS
carc_CNS_box_baseline = ggplot(cvi_carc_CNS, aes(x = chemical, y = Baseline.Vulnerabilities)) + geom_boxplot(position = position_dodge()) + theme_bw() + ylab("Baseline Vulnerability") + xlab("Chemical")  + theme(axis.text.x = element_text(size = 10, angle = 45, hjust = 1)) + scale_y_continuous(breaks = seq(0.2, 0.8, 0.1), limits = c(0.2,0.8))

carc_CNS_box_baseline

ggsave(filename = "carc_CNS_box_baseline.png", device = "png", plot = carc_CNS_box_baseline, path = "~/TSCA/CRA/CRA_Prioritization/Output/", width = 10, height = 5)

#cardiorespiratory
cardio_box_baseline = ggplot(cvi_cardio, aes(x = chemical, y = Baseline.Vulnerabilities)) + geom_boxplot(position = position_dodge()) + theme_bw() + ylab("Baseline Vulnerability") + xlab("Chemical")  + theme(axis.text.x = element_text(size = 10, angle = 45, hjust = 1)) + scale_y_continuous(breaks = seq(0.2, 0.8, 0.1), limits = c(0.2,0.8))

cardio_box_baseline

ggsave(filename = "cardio_box_baseline.png", device = "png", plot = cardio_box_baseline, path = "~/TSCA/CRA/CRA_Prioritization/Output/", width = 10, height = 5)

#EDCs
edcs_box_baseline = ggplot(cvi_edcs, aes(x = chemical, y = Baseline.Vulnerabilities)) + geom_boxplot(position = position_dodge()) + theme_bw() + ylab("Baseline Vulnerability") + xlab("Chemical")  + theme(axis.text.x = element_text(size = 10, angle = 45, hjust = 1)) + scale_y_continuous(breaks = seq(0.2, 0.8, 0.1), limits = c(0.2,0.8))

edcs_box_baseline

ggsave(filename = "edcs_box_baseline.png", device = "png", plot = edcs_box_baseline, path = "~/TSCA/CRA/CRA_Prioritization/Output/", width = 10, height = 5)

#liver, kidney, thyroid
lkt_box_baseline = ggplot(cvi_lkt, aes(x = chemical, y = Baseline.Vulnerabilities)) + geom_boxplot(position = position_dodge()) + theme_bw() + ylab("Baseline Vulnerability") + xlab("Chemical")  + theme(axis.text.x = element_text(size = 10, angle = 45, hjust = 1)) + scale_y_continuous(breaks = seq(0.2, 0.8, 0.1), limits = c(0.2,0.8))

lkt_box_baseline

ggsave(filename = "lkt_box_baseline.png", device = "png", plot = lkt_box_baseline, path = "~/TSCA/CRA/CRA_Prioritization/Output/", width = 10, height = 5)
```


#repeat everything we did with baseline for climate vulnerability index to see where climate might have an impact (want to do just for climate or overall cvi? doing for overall cvi for now)


#Get average CVI score for each chemical and endpoint group
```{r}
cvi_bychem = summarySE(tri_cvi, groupvars = "chemical", measurevar = "CVI.Score")

#CNS chemicals
cvi_CNS = filter(cvi_bychem, chemical == "Cyanide compounds" | chemical == "Hydrogen cyanide" |  chemical == "Ethylbenzene" | chemical == "m-Xylene" | chemical == "o-Xylene" | chemical == "p-Xylene" | chemical == "Xylene (mixed isomers)")

#Carcinogens with CNS
cvi_carc_CNS = filter(cvi_bychem, chemical == "Acetaldehyde" | chemical == "Acrylonitrile"|chemical == "Aniline"|chemical == "Styrene"|chemical == "Benzene"|chemical == "Bromoform"|chemical == "Vinyl chloride")

#Carcinogens
cvi_carc = filter(cvi_bychem, chemical == "Creosote"|chemical == "3,3'-Dichlorobenzidine"|chemical == "3,3'-Dichlorobenzidine dihydrochloride"|chemical == "4,4'-Methylenebis(2-chloroaniline)"|chemical == "N-Nitrosodiphenylamine")

#Carcinogenetic metals
cvi_carc_metals = filter(cvi_bychem, chemical == "Cadmium" | chemical == "Cadmium compounds"|chemical == "Nickel" | chemical == "Nickel compounds"|chemical == "Arsenic" | chemical == "Arsenic compounds"|chemical == "Chromium" | chemical == "Chromium compounds"|chemical == "Lead" | chemical == "Lead compounds")

#cardiorespiratory
cvi_cardio = filter(cvi_bychem, chemical == "Barium" | chemical == "Barium compounds"|chemical == "Cobalt" | chemical == "Cobalt compounds"|chemical == "Molybdenum trioxide"|chemical == "Naphthalene"|chemical == "Maleic anhydride"|chemical == "1-Hexadecanol"|chemical == "Hydrogen fluoride") 

#liver, kidney, thyroid
cvi_lkt = filter(cvi_bychem, chemical == "Hexachloro-1,3-butadiene")

#edcs
cvi_edcs = filter(cvi_bychem, chemical == "Nonylphenol" | chemical == "Nonylphenol Ethoxylates"|chemical == "4,4'-Isopropylidenediphenol"|chemical == "Decabromodiphenyl oxide")

```


#add groupings to climate vulnerability index dfs so I can add back into one big df
```{r}
cvi_carc$endpoint = "carc"
cvi_carc_CNS$endpoint = "carc_CNS"
cvi_carc_metals$endpoint = "carc_metals"
cvi_cardio$endpoint = "cardio"
cvi_CNS$endpoint = "CNS"
cvi_edcs$endpoint = "edcs"
cvi_lkt$endpoint = "lkt"

all_cvi = rbind(cvi_carc, cvi_carc_CNS, cvi_carc_metals, cvi_cardio, cvi_CNS, cvi_edcs, cvi_lkt)

tri_cvi = merge(tri_cvi, all_cvi[,c("chemical", "endpoint")], by = "chemical")
```


#summary and stats to compare between end groups and graph for climate vulnerability index
```{r}
tri_cvi_aov = aov(CVI.Score ~ endpoint, data = tri_cvi)
tri_cvi_aov
tri_cvi_tukey = TukeyHSD(x = tri_cvi_aov, conf.level = 0.95)
tri_cvi_tukey
capture.output(tri_cvi_tukey, file = "~/TSCA/CRA/CRA_Prioritization/Output/tri_cvi_tukey")

#Boxplot 
endpoints = c("Carcinogens", "Carcinogenic CNS", "Carcinogenetic Metals", "Cardiorespiratory", "CNS", "EDCs", "Liver, Kidney, Thyroid")

tri_cvi_box = ggplot(tri_cvi, aes(x = endpoint, y = CVI.Score)) + geom_boxplot(position = position_dodge()) + theme_bw() + ylab("CVI Score") + xlab("Endpoint Group") + scale_x_discrete(labels = endpoints) + theme(axis.text.x = element_text(size = 10, angle = 45, hjust = 1))

tri_cvi_box

ggsave(filename = "tri_cvi_box.png", device = "png", plot = tri_cvi_box, path = "~/TSCA/CRA/CRA_Prioritization/Output/", width = 10, height = 5)

```


#filter tri_cvi for each endpoint group (without summary numbers, like above in cvi_endpoint dfs)
```{r}
cvi_CNS = filter(tri_cvi, chemical == "Cyanide compounds" | chemical == "Hydrogen cyanide" |  chemical == "Ethylbenzene" | chemical == "m-Xylene" | chemical == "o-Xylene" | chemical == "p-Xylene" | chemical == "Xylene (mixed isomers)")

cvi_carc_CNS = filter(tri_cvi, chemical == "Acetaldehyde" | chemical == "Acrylonitrile"|chemical == "Aniline"|chemical == "Styrene"|chemical == "Benzene"|chemical == "Bromoform"|chemical == "Vinyl chloride")

cvi_carc = filter(tri_cvi, chemical == "Creosote"|chemical == "3,3'-Dichlorobenzidine"|chemical == "3,3'-Dichlorobenzidine dihydrochloride"|chemical == "4,4'-Methylenebis(2-chloroaniline)"|chemical == "N-Nitrosodiphenylamine")

cvi_carc_metals = filter(tri_cvi, chemical == "Cadmium" | chemical == "Cadmium compounds"|chemical == "Nickel" | chemical == "Nickel compounds"|chemical == "Arsenic" | chemical == "Arsenic compounds"|chemical == "Chromium" | chemical == "Chromium compounds"|chemical == "Lead" | chemical == "Lead compounds")

cvi_cardio = filter(tri_cvi, chemical == "Barium" | chemical == "Barium compounds"|chemical == "Cobalt" | chemical == "Cobalt compounds"|chemical == "Molybdenum trioxide"|chemical == "Naphthalene"|chemical == "Maleic anhydride"|chemical == "1-Hexadecanol"|chemical == "Hydrogen fluoride") 

cvi_lkt = filter(tri_cvi, chemical == "Hexachloro-1,3-butadiene")

cvi_edcs = filter(tri_cvi, chemical == "Nonylphenol" | chemical == "Nonylphenol Ethoxylates"|chemical == "4,4'-Isopropylidenediphenol"|chemical == "Decabromodiphenyl oxide")
```


#boxplots for climate vulnerability of each chemical within endpoint groups
```{r}
#carcinogens
carc_box_cvi = ggplot(cvi_carc, aes(x = chemical, y = CVI.Score)) + geom_boxplot(position = position_dodge()) + theme_bw() + ylab("CVI Score") + xlab("Chemical")  + theme(axis.text.x = element_text(size = 10, angle = 45, hjust = 1)) + scale_y_continuous(breaks = seq(0.2, 0.8, 0.1), limits = c(0.2,0.8))

carc_box_cvi

ggsave(filename = "carc_box_cvi.png", device = "png", plot = carc_box_cvi, path = "~/TSCA/CRA/CRA_Prioritization/Output/", width = 10, height = 5)

#CNS
CNS_box_cvi = ggplot(cvi_CNS, aes(x = chemical, y = CVI.Score)) + geom_boxplot(position = position_dodge()) + theme_bw() + ylab("CVI Score") + xlab("Chemical")  + theme(axis.text.x = element_text(size = 10, angle = 45, hjust = 1)) + scale_y_continuous(breaks = seq(0.2, 0.8, 0.1), limits = c(0.2,0.8))

CNS_box_cvi

ggsave(filename = "CNS_box_cvi.png", device = "png", plot = CNS_box_cvi, path = "~/TSCA/CRA/CRA_Prioritization/Output/", width = 10, height = 5)

#carcinogenic metals
carc_metals_box_cvi = ggplot(cvi_carc_metals, aes(x = chemical, y = CVI.Score)) + geom_boxplot(position = position_dodge()) + theme_bw() + ylab("CVI Score") + xlab("Chemical")  + theme(axis.text.x = element_text(size = 10, angle = 45, hjust = 1)) + scale_y_continuous(breaks = seq(0.2, 0.8, 0.1), limits = c(0.2,0.8))

carc_metals_box_cvi

ggsave(filename = "carc_metals_box_cvi.png", device = "png", plot = carc_metals_box_cvi, path = "~/TSCA/CRA/CRA_Prioritization/Output/", width = 10, height = 5)

#carcinogenic with CNS
carc_CNS_box_cvi = ggplot(cvi_carc_CNS, aes(x = chemical, y = CVI.Score)) + geom_boxplot(position = position_dodge()) + theme_bw() + ylab("CVI Score") + xlab("Chemical")  + theme(axis.text.x = element_text(size = 10, angle = 45, hjust = 1)) + scale_y_continuous(breaks = seq(0.2, 0.8, 0.1), limits = c(0.2,0.8))

carc_CNS_box_cvi

ggsave(filename = "carc_CNS_box_cvi.png", device = "png", plot = carc_CNS_box_cvi, path = "~/TSCA/CRA/CRA_Prioritization/Output/", width = 10, height = 5)

#cardiorespiratory
cardio_box_cvi = ggplot(cvi_cardio, aes(x = chemical, y = CVI.Score)) + geom_boxplot(position = position_dodge()) + theme_bw() + ylab("CVI Score") + xlab("Chemical")  + theme(axis.text.x = element_text(size = 10, angle = 45, hjust = 1)) + scale_y_continuous(breaks = seq(0.2, 0.8, 0.1), limits = c(0.2,0.8))

cardio_box_cvi

ggsave(filename = "cardio_box_cvi.png", device = "png", plot = cardio_box_cvi, path = "~/TSCA/CRA/CRA_Prioritization/Output/", width = 10, height = 5)

#EDCs
edcs_box_cvi = ggplot(cvi_edcs, aes(x = chemical, y = CVI.Score)) + geom_boxplot(position = position_dodge()) + theme_bw() + ylab("CVI Score") + xlab("Chemical")  + theme(axis.text.x = element_text(size = 10, angle = 45, hjust = 1)) + scale_y_continuous(breaks = seq(0.2, 0.8, 0.1), limits = c(0.2,0.8))

edcs_box_cvi

ggsave(filename = "edcs_box_cvi.png", device = "png", plot = edcs_box_cvi, path = "~/TSCA/CRA/CRA_Prioritization/Output/", width = 10, height = 5)

#liver, kidney, thyroid
lkt_box_cvi = ggplot(cvi_lkt, aes(x = chemical, y = CVI.Score)) + geom_boxplot(position = position_dodge()) + theme_bw() + ylab("CVI Score") + xlab("Chemical")  + theme(axis.text.x = element_text(size = 10, angle = 45, hjust = 1)) + scale_y_continuous(breaks = seq(0.2, 0.8, 0.1), limits = c(0.2,0.8))

lkt_box_cvi

ggsave(filename = "lkt_box_cvi.png", device = "png", plot = lkt_box_cvi, path = "~/TSCA/CRA/CRA_Prioritization/Output/", width = 10, height = 5)
```









