---
title: "CVI"
author: "Paige Varner"
date: "2024-09-17"
output: html_document
---

#load libraries
```{r}
library("dplyr")
library("bigrquery")
library("DBI")
library("pheatmap")
library("RColorBrewer")
library("Rmisc")
library("stats")
library("rstatix")
library("ggplot2")
library("tigris")
library("sf")
library("tidycensus")
library("viridis")
library("tidyr")
library("car")
library("pak")
library("magick")

#install patch for Tigris as of April 2025
#pak::pak("walkerke/tigris")

detach(package:Rmisc)
detach(package:plyr)
```


#query data from BQ (dataframe with TRI ID, facility name, chemicals, lat/long, and total sum of on site releases for the 5 year span)

```{r}
	# set up database connection
	con <- dbConnect(
	  bigrquery::bigquery(),
	  project = "edf-aq-data",
	  dataset = "tri",
)

SQL = "	
	SELECT
	trifd,
	facility_name,
	chemical,
	latitude,
	longitude,
	SUM(on_site_release_total) as sum_onsite_releases_2017_2022
	FROM `edf-aq-data.tri.tri_basic_2017-2022` 
	GROUP BY trifd, facility_name, chemical, latitude, longitude"

tri = dbGetQuery(con,SQL)

tri = filter(tri, sum_onsite_releases_2017_2022 > 0)
```


#fix multiple chemical names
```{r}
tri$chemical = gsub("Hydrogen cyanide", "Cyanide compounds", tri$chemical)
tri$chemical = gsub("Xylene \\(mixed isomers\\)", "Xylenes", tri$chemical)
tri$chemical = gsub("o-Xylene", "Xylenes", tri$chemical)
tri$chemical = gsub("p-Xylene", "Xylenes", tri$chemical)
tri$chemical = gsub("m-Xylene", "Xylenes", tri$chemical)
tri$chemical = gsub("Cadmium compounds", "Cadmium", tri$chemical)
tri$chemical = gsub("Arsenic compounds", "Arsenic", tri$chemical)
tri$chemical = gsub("Nickel compounds", "Nickel", tri$chemical)
tri$chemical = gsub("Lead compounds", "Lead", tri$chemical)
tri$chemical = gsub("Barium compounds", "Barium", tri$chemical)
tri$chemical = gsub("Cobalt compounds", "Cobalt", tri$chemical)
tri$chemical = gsub("Chromium compounds", "Chromium", tri$chemical)
tri$chemical = gsub("Nonylphenol Ethoxylates", "Nonylphenol", tri$chemical)


tri = tri[!duplicated(tri[c("trifd", "chemical")]),]

#add up chemicals with same name for each trifd to get total releases of those chemicals 
tri <- tri %>%
  group_by(trifd, latitude, longitude, chemical) %>%
  summarise(onsite_releases_2017_2022 = sum(sum_onsite_releases_2017_2022, na.rm = TRUE))

```


#get census tract GEOIDs for all TRI facilities based on long/lat (from 2022 census)
```{r}
cvi = read.csv("~/TSCA/CRA/CRA_Prioritization/Data/cvi_scores.csv")
colnames(cvi)[1] = "GEOID"

#getting census tract info for all TRI facilities
tracts = tracts(state = NULL, cb = TRUE)

tri.points <- tri %>%
  filter(!is.na(latitude)) %>%  # remove missing
  st_as_sf(coords=c('longitude', 'latitude'), crs=st_crs(tracts))

tri.tracts = st_join(tri.points, tracts)
```


#join CVI data to TRI data based on GEOID
```{r}
tri_cvi = merge(tri.tracts, cvi, by = "GEOID")
write.csv(tri_cvi, "~/TSCA/CRA/CRA_Prioritization/Data/tri_cvi_2017_2022.csv")
```


#add endpoint designations for each chemical and filter all TRI and CVI info by endpoints
```{r}
endpoints = read.csv("~/TSCA/CRA/CRA_Prioritization/Data/TSCA_endpoints.csv")

tri_cvi_endpoints <- tri_cvi %>%
  left_join(endpoints, by = c("chemical" = "tri_name")) %>%
  filter(!is.na(analysis_name)) %>%
  select(-analysis_name, -chemical.y)

carc <- tri_cvi_endpoints %>%
  filter(carc %in% c("Known", "Probable"))

cns <- tri_cvi_endpoints %>%
  filter(cns %in% "yes")

cardioresp = tri_cvi_endpoints %>%
  filter(cardio == "yes" | resp == "yes")

devrep = tri_cvi_endpoints %>%
  filter(dev == "yes" | repro == "yes")

lkt = tri_cvi_endpoints %>%
  filter(lkt %in% "yes")
```


#summary and stats to compare CVI score between endpoint groups
```{r}
#Create a long format dataframe where each row can belong to multiple endpoints
tri_cvi_endpoints_long <- tri_cvi_endpoints %>%
  mutate(
    carc_group = ifelse(carc %in% c("Known", "Probable"), "carc", NA_character_),
    cns_group = ifelse(cns == "yes", "cns", NA_character_),
    devrep_group = ifelse(dev == "yes" | repro == "yes", "devrep", NA_character_),
    cardioresp_group = ifelse(resp == "yes" | cardio == "yes", "cardioresp", NA_character_),
    lkt_group = ifelse(lkt == "yes", "lkt", NA_character_)
  ) %>%
  pivot_longer(cols = carc_group:lkt_group, names_to = "endpoint_type", values_to = "endpoint_group") %>%
  filter(!is.na(endpoint_group))

#statistical analysis
cvi_aov = aov(CVI.Score ~ endpoint_group, data = tri_cvi_endpoints_long)
summary(cvi_aov)

cvi_tukey = TukeyHSD(x = cvi_aov, conf.level = 0.95)
cvi_tukey

capture.output(cvi_tukey, file = "~/TSCA/CRA/CRA_Prioritization/Output/cvi_tukey")

#summary stats within endpoints for reporting in table
CVI_summary = tri_cvi_endpoints_long %>%
  group_by(endpoint_group) %>%
  summarize(median_CVI = median(CVI.Score, na.rm = TRUE),
  mean_CVI = mean(CVI.Score, na.rm = TRUE),      
  sd_CVI = sd(CVI.Score, na.rm = TRUE),   
  min_CVI = min(CVI.Score, na.rm = TRUE),
  max_CVI = max(CVI.Score, na.rm = TRUE),
  n = n(),                                       
  se_CVI = sd_CVI / sqrt(n), 
  .groups = "drop") %>%
  st_drop_geometry()

write.csv(CVI_summary, file = "~/TSCA/CRA/CRA_Prioritization/Output/cvi_summary.csv")
```



#get average CVI score for each chemical within endpoint groups for reporting purposes 
```{r}
carc_CVI_summary = carc %>%
  group_by(chemical) %>%
  summarize(
    median_CVI = median(CVI.Score, na.rm = TRUE),
    mean_CVI = mean(CVI.Score, na.rm = TRUE),      
    sd_CVI = sd(CVI.Score, na.rm = TRUE),  
    min_CVI = min(CVI.Score, na.rm = TRUE),
    max_CVI = max(CVI.Score, na.rm = TRUE),        
    n = n(),                                       
    se_CVI = sd_CVI / sqrt(n), 
    .groups = "drop"
  ) %>%
  st_drop_geometry()

cns_CVI_summary = cns %>%
  group_by(chemical) %>%
  summarize(
    median_CVI = median(CVI.Score, na.rm = TRUE),
    mean_CVI = mean(CVI.Score, na.rm = TRUE),      
    sd_CVI = sd(CVI.Score, na.rm = TRUE),  
    min_CVI = min(CVI.Score, na.rm = TRUE),
    max_CVI = max(CVI.Score, na.rm = TRUE),        
    n = n(),                                       
    se_CVI = sd_CVI / sqrt(n), 
    .groups = "drop"
  ) %>%
  st_drop_geometry()

cardioresp_CVI_summary = cardioresp %>%
  group_by(chemical) %>%
  summarize(
    median_CVI = median(CVI.Score, na.rm = TRUE),
    mean_CVI = mean(CVI.Score, na.rm = TRUE),      
    sd_CVI = sd(CVI.Score, na.rm = TRUE),  
    min_CVI = min(CVI.Score, na.rm = TRUE),
    max_CVI = max(CVI.Score, na.rm = TRUE),        
    n = n(),                                       
    se_CVI = sd_CVI / sqrt(n), 
    .groups = "drop"
  ) %>%
  st_drop_geometry()

devrep_CVI_summary = devrep %>%
  group_by(chemical) %>%
  summarize(
    median_CVI = median(CVI.Score, na.rm = TRUE),
    mean_CVI = mean(CVI.Score, na.rm = TRUE),      
    sd_CVI = sd(CVI.Score, na.rm = TRUE),  
    min_CVI = min(CVI.Score, na.rm = TRUE),
    max_CVI = max(CVI.Score, na.rm = TRUE),        
    n = n(),                                       
    se_CVI = sd_CVI / sqrt(n), 
    .groups = "drop"
  ) %>%
  st_drop_geometry()

lkt_CVI_summary = lkt %>%
  group_by(chemical) %>%
  summarize(
    median_CVI = median(CVI.Score, na.rm = TRUE),
    mean_CVI = mean(CVI.Score, na.rm = TRUE),      
    sd_CVI = sd(CVI.Score, na.rm = TRUE),  
    min_CVI = min(CVI.Score, na.rm = TRUE),
    max_CVI = max(CVI.Score, na.rm = TRUE),        
    n = n(),                                       
    se_CVI = sd_CVI / sqrt(n), 
    .groups = "drop"
  ) %>%
  st_drop_geometry()


write.csv(carc_CVI_summary, file = "~/TSCA/CRA/CRA_Prioritization/Output/carc_cvi_summary.csv")
write.csv(cns_CVI_summary, file = "~/TSCA/CRA/CRA_Prioritization/Output/cns_cvi_summary.csv")
write.csv(cardioresp_CVI_summary, file = "~/TSCA/CRA/CRA_Prioritization/Output/cardioresp_cvi_summary.csv")
write.csv(devrep_CVI_summary, file = "~/TSCA/CRA/CRA_Prioritization/Output/devrep_cvi_summary.csv")
write.csv(lkt_CVI_summary, file = "~/TSCA/CRA/CRA_Prioritization/Output/lkt_cvi_summary.csv")
```


#edit names of some chemicals for graphing
```{r}
carc <- carc %>%
    mutate(
    chemical = ifelse(chemical == "4,4'-Methylenebis(2-chloroaniline)", "MBOCA", chemical),
    chemical = ifelse(chemical == "3,3'-Dichlorobenzidine dihydrochloride", "3,3'-Dichlorobenzidine 2HCl", chemical)
  ) 

cns = cns %>%
  mutate(chemical = ifelse(chemical == "Decabromodiphenyl oxide", "DecaBDE", chemical), chemical = ifelse(chemical == "Cyanide compounds", "Cyanide", chemical))

cardioresp = cardioresp %>%
  mutate(chemical = ifelse(chemical == "Molybdenum trioxide", "Molybdenum", chemical))

devrep = devrep %>%
  mutate(chemical = ifelse(chemical == "Molybdenum trioxide", "Molybdenum", chemical),chemical = ifelse(chemical == "4,4'-Isopropylidenediphenol", "Bisphenol A", chemical),chemical = ifelse(chemical == "Decabromodiphenyl oxide", "DecaBDE", chemical), chemical = ifelse(chemical == "Hexachloro-1,3-butadiene", "Hexachlorobutadiene", chemical))

lkt = lkt %>%
  mutate(chemical1 = ifelse(chemical == "Molybdenum trioxide", "Molybdenum", chemical),chemical = ifelse(chemical == "Decabromodiphenyl oxide", "DecaBDE", chemical),chemical = ifelse(chemical == "4,4'-Methylenebis(2-chloroaniline)", "MBOCA", chemical), chemical = ifelse(chemical == "Hexachloro-1,3-butadiene", "Hexachlorobutadiene", chemical))
```



#violin for CVI score of each chemical within endpoint groups
```{r}
## carcinogens

# Compute sample size for each chemical
chemical_counts <- carc %>%
  group_by(chemical) %>%
  summarise(sample_size = n())

chemical_labels <- setNames(
  paste0(chemical_counts$chemical, " (", chemical_counts$sample_size, ")"),
  chemical_counts$chemical
)

#plot
carc_violin <- ggplot(carc, aes(x = chemical, y = CVI.Score, fill = chemical)) +
  geom_violin(trim = FALSE) +
  theme_bw() +
  ylab("CVI Score") +
  xlab("Chemical") +
  theme(
    axis.text.x = element_text(size = 14, angle = 60, hjust = 1), 
    axis.text.y = element_text(size = 14),  
    axis.title.x = element_text(size = 16, face = "bold"),  
    axis.title.y = element_text(size = 16, face = "bold"),  
    legend.position = "none"
  ) +
  stat_summary(fun = mean, geom = "point", shape = 20, size = 5, 
               color = "black", fill = "gray") +
  geom_hline(yintercept = mean(carc$CVI.Score), linetype = "dashed") +
  scale_y_continuous(breaks = seq(0.2, 0.8, 0.1), limits = c(0.2, 0.8)) +
  scale_fill_viridis(discrete = TRUE, option = "H", begin = 0.2) +
  scale_x_discrete(labels = chemical_labels)  

carc_violin

ggsave(filename = "carc_violin.png", device = "png", plot = carc_violin, path = "~/TSCA/CRA/CRA_Prioritization/Output/", width = 13, height = 5)

## CNS

# Compute sample size for each chemical
chemical_counts <- cns %>%
  group_by(chemical) %>%
  summarise(sample_size = n())

chemical_labels <- setNames(
  paste0(chemical_counts$chemical, " (", chemical_counts$sample_size, ")"),
  chemical_counts$chemical
)

# Plot
cns_violin <- ggplot(cns, aes(x = chemical, y = CVI.Score, fill = chemical)) +
  geom_violin(trim = FALSE) +
  theme_bw() +
  ylab("CVI Score") +
  xlab("Chemical") +
  theme(
    axis.text.x = element_text(size = 14, angle = 55, hjust = 1), 
    axis.text.y = element_text(size = 14),  
    axis.title.x = element_text(size = 16, face = "bold"),  
    axis.title.y = element_text(size = 16, face = "bold"),  
    legend.position = "none"
  ) +
  stat_summary(fun = mean, geom = "point", shape = 20, size = 5, 
               color = "black", fill = "gray") +
  geom_hline(yintercept = mean(cns$CVI.Score), linetype = "dashed") +
  scale_y_continuous(breaks = seq(0.2, 0.8, 0.1), limits = c(0.2, 0.8)) +
  scale_fill_viridis(discrete = TRUE, option = "H", begin = 0.2) +
  scale_x_discrete(labels = chemical_labels) 

cns_violin

ggsave(filename = "cns_violin.png", device = "png", plot = cns_violin, 
       path = "~/TSCA/CRA/CRA_Prioritization/Output/", width = 11, height = 5)


## cardiorespiratory

# Compute sample size for each chemical
chemical_counts <- cardioresp %>%
  group_by(chemical) %>%
  summarise(sample_size = n())

chemical_labels <- setNames(
  paste0(chemical_counts$chemical, " (", chemical_counts$sample_size, ")"),
  chemical_counts$chemical
)

# Plot
cardioresp_violin <- ggplot(cardioresp, aes(x = chemical, y = CVI.Score, fill = chemical)) +
  geom_violin(trim = FALSE) +
  theme_bw() +
  ylab("CVI Score") +
  xlab("Chemical") +
  theme(
    axis.text.x = element_text(size = 14, angle = 55, hjust = 1), 
    axis.text.y = element_text(size = 14),  
    axis.title.x = element_text(size = 16, face = "bold"),  
    axis.title.y = element_text(size = 16, face = "bold"),  
    legend.position = "none"
  ) +
  stat_summary(fun = mean, geom = "point", shape = 20, size = 5, 
               color = "black", fill = "gray") +
  geom_hline(yintercept = mean(cardioresp$CVI.Score), linetype = "dashed") +
  scale_y_continuous(breaks = seq(0.2, 0.8, 0.1), limits = c(0.2, 0.8)) +
  scale_fill_viridis(discrete = TRUE, option = "H", begin = 0.2) +
  scale_x_discrete(labels = chemical_labels) 

cardioresp_violin

ggsave(filename = "cardioresp_violin.png", device = "png", plot = cardioresp_violin, 
       path = "~/TSCA/CRA/CRA_Prioritization/Output/", width = 14, height = 5)

## dev/rep

# Compute sample size for each chemical
chemical_counts <- devrep %>%
  group_by(chemical) %>%
  summarise(sample_size = n())

chemical_labels <- setNames(
  paste0(chemical_counts$chemical, " (", chemical_counts$sample_size, ")"),
  chemical_counts$chemical
)

# Plot
devrep_violin <- ggplot(devrep, aes(x = chemical, y = CVI.Score, fill = chemical)) +
  geom_violin(trim = FALSE) +
  theme_bw() +
  ylab("CVI Score") +
  xlab("Chemical") +
  theme(
    axis.text.x = element_text(size = 14, angle = 55, hjust = 1), 
    axis.text.y = element_text(size = 14),  
    axis.title.x = element_text(size = 16, face = "bold"),  
    axis.title.y = element_text(size = 16, face = "bold"),  
    legend.position = "none"
  ) +
  stat_summary(fun = mean, geom = "point", shape = 20, size = 5, 
               color = "black", fill = "gray") +
  geom_hline(yintercept = mean(devrep$CVI.Score), linetype = "dashed") +
  scale_y_continuous(breaks = seq(0.2, 0.8, 0.1), limits = c(0.2, 0.8)) +
  scale_fill_viridis(discrete = TRUE, option = "H", begin = 0.2) +
  scale_x_discrete(labels = chemical_labels) 

devrep_violin

ggsave(filename = "devrep_violin.png", device = "png", plot = devrep_violin, 
       path = "~/TSCA/CRA/CRA_Prioritization/Output/", width = 12, height = 5)


## liver, kidney, thyroid

# Compute sample size for each chemical
chemical_counts <- lkt %>%
  group_by(chemical) %>%
  summarise(sample_size = n())

chemical_labels <- setNames(
  paste0(chemical_counts$chemical, " (", chemical_counts$sample_size, ")"),
  chemical_counts$chemical
)

# Plot
lkt_violin <- ggplot(lkt, aes(x = chemical, y = CVI.Score, fill = chemical)) +
  geom_violin(trim = FALSE) +
  theme_bw() +
  ylab("CVI Score") +
  xlab("Chemical") +
  theme(
    axis.text.x = element_text(size = 14, angle = 55, hjust = 1), 
    axis.text.y = element_text(size = 14),  
    axis.title.x = element_text(size = 16, face = "bold"),  
    axis.title.y = element_text(size = 16, face = "bold"),  
    legend.position = "none"
  ) +
  stat_summary(fun = mean, geom = "point", shape = 20, size = 5, 
               color = "black", fill = "gray") +
  geom_hline(yintercept = mean(lkt$CVI.Score), linetype = "dashed") +
  scale_y_continuous(breaks = seq(0.2, 0.8, 0.1), limits = c(0.2, 0.8)) +
  scale_fill_viridis(discrete = TRUE, option = "H", begin = 0.2) +
  scale_x_discrete(labels = chemical_labels) 

lkt_violin

ggsave(filename = "lkt_violin.png", device = "png", plot = lkt_violin, 
       path = "~/TSCA/CRA/CRA_Prioritization/Output/", width = 12, height = 5)



```




#stats for above: significant differences between CVI scores of chemicals in each endpoint group
```{r}
## carc
carc_cvi_aov = aov(CVI.Score ~ chemical, data = carc)
summary(carc_cvi_aov)

#Shapiro-Wilk test for normality (p < 0.05, data is not normally distributed; histogram looks SUPER normally distributed, though?)
shapiro.test(carc_cvi_aov$residuals[0:5000])
hist(carc$CVI.Score)

#Levene's Test for homogeneity of variances (p < 0.05, variances are not homogenous)
leveneTest(CVI.Score ~ chemical, data = carc)

#log-transform and check distribution with histogram and tests (made data more skewed and barely reduced diff between variances; not going to log transform)
carc$log_CVI.Score <- log(carc$CVI.Score)
hist(carc$log_CVI.Score)
leveneTest(log_CVI.Score ~ chemical, data = carc) 

carc_cvi_tukey = TukeyHSD(x = carc_cvi_aov, conf.level = 0.95)
capture.output(carc_cvi_tukey, file = "~/TSCA/CRA/CRA_Prioritization/Output/carc_cvi_tukey")

#cns
cns_cvi_aov = aov(CVI.Score ~ chemical, data = cns)
summary(cns_cvi_aov)
cns_cvi_tukey = TukeyHSD(x = cns_cvi_aov, conf.level = 0.95)
capture.output(cns_cvi_tukey, file = "~/TSCA/CRA/CRA_Prioritization/Output/cns_cvi_tukey")

#cardioresp
cardioresp_cvi_aov = aov(CVI.Score ~ chemical, data = cardioresp)
summary(cardioresp_cvi_aov)
cardioresp_cvi_tukey = TukeyHSD(x = cardioresp_cvi_aov, conf.level = 0.95)
capture.output(cardioresp_cvi_tukey, file = "~/TSCA/CRA/CRA_Prioritization/Output/cardioresp_cvi_tukey")

#devrep
devrep_cvi_aov = aov(CVI.Score ~ chemical, data = devrep)
summary(devrep_cvi_aov)
devrep_cvi_tukey = TukeyHSD(x = devrep_cvi_aov, conf.level = 0.95)
capture.output(devrep_cvi_tukey, file = "~/TSCA/CRA/CRA_Prioritization/Output/devrep_cvi_tukey")

#lkt
lkt_cvi_aov = aov(CVI.Score ~ chemical, data = lkt)
summary(lkt_cvi_aov)
lkt_cvi_tukey = TukeyHSD(x = lkt_cvi_aov, conf.level = 0.95)
capture.output(lkt_cvi_tukey, file = "~/TSCA/CRA/CRA_Prioritization/Output/lkt_cvi_tukey")
```



#filter for most vulnerable census tracts by baseline vulnerability
```{r}
#baseline vulnerability 
cutoff = quantile(tri_cvi_endpoints$Baseline.Vulnerabilities, 0.75, na.rm = TRUE)

tri_cvi_25thperc = tri_cvi_endpoints %>%
  filter(CVI.Score > cutoff)

write.csv(tri_cvi_25thperc, file = "~/TSCA/CRA/CRA_Prioritization/Output/tri_cvi_25thperc_baseline.csv")
```


#sum up releases per chemical to see which chemicals are released in the highest amounts in the top 75th - 100th percentile of baseline vulnerability
```{r}

baseline_sum <- tri_cvi_25thperc %>%
  group_by(chemical) %>%
  summarise(
    total_onsite_releases = sum(onsite_releases_2017_2022, na.rm = TRUE),
    sample_size = n()
  )

baseline_sum = st_drop_geometry(baseline_sum)

write.csv(baseline_sum, file = "~/TSCA/CRA/CRA_Prioritization/Output/25thperc_baseline_sum_allchems.csv")
```



#combine images for MS
```{r}
# Set working directory
setwd("~/TSCA/CRA/CRA_Prioritization/Output/")

# Define file names and labels
files <- c("carc_violin.png", "cns_violin.png", "cardioresp_violin.png", "devrep_violin.png", "lkt_violin.png")
labels <- c("A", "B", "C", "D", "E")

# Determine max image width
max_width <- max(sapply(files, function(f) image_info(image_read(f))$width))

# Set label image height and font size
label_height <- 100
label_font_size <- 80

# Create labeled image list
imgs <- mapply(function(file, label) {
  # Read and prepare the image
  img <- image_read(file)
  img <- image_background(img, "white", flatten = TRUE)
  img <- image_resize(img, paste0(max_width))
  
  # Create a white label strip
  label_img <- image_blank(width = max_width, height = label_height, color = "white")
  
  # Left-align label in the strip
  label_img <- image_annotate(
    label_img, label,
    size = label_font_size,
    gravity = "west",          # Align to left side
    location = "+20+0",        # Slight right padding
    color = "black"
  )
  
  # Stack label above the image
  image_append(c(label_img, img), stack = TRUE)
}, files, labels, SIMPLIFY = FALSE)

# Combine all vertically
combined <- image_append(image_join(imgs), stack = TRUE)

# Ensure white background
combined <- image_background(combined, "white", flatten = TRUE)

# Save as high-resolution PDF
image_write(combined, path = "combined_violin.pdf", density = "300x300", format = "pdf")


```






